### 2014-12-02 proposal.tex
proposal

\documentclass[12pt]{article}

\usepackage{amsmath}    % just math
\usepackage{amssymb}    % allow blackboard bold (aka N,R,Q sets)
\usepackage{amsthm} % allows thm environment
\usepackage{graphicx, framed} % allows graphics
\usepackage[none]{hyphenat} % allows graphics
%\usepackage[margin=1.5cm, vmargin={0pt, 1cm}, includefoot]{geometry}

\title{STA206 Project Proposal \\ Predicting Exchange Rates}
\author{Clark Fitzgerald, Amy Kim}

\usepackage[usenames,dvipsnames]{color}
\usepackage{color}
\usepackage{wrapfig}
\usepackage{ulem}

\linespread{1}  % single spaces lines

\textwidth 6.5truein  % These 4 commands define more efficient margins
\textheight 9.5truein
\oddsidemargin 0.0in
\topmargin -0.6in

\parskip 5pt  % Also, a bit of space between paragraphs

\begin{document}
\maketitle
\paragraph{1. Group Membership}
\begin{itemize}
    \item Clark Fitzgerald (the contact person) clarkfitzg@gmail.com
    \item Amy Kim atykim@ucdavis.edu
\end{itemize}

\paragraph{2. Which Data to Analyze}

We use data from Quandl.com, a service that provides clean, documented
data from a wide variety of sources. The data is loaded directly using a
web API.
Using this service will make it easier
to repeat the analysis in the future, or conduct similar analyses between
different countries.

The number of variable set in our data is 22 (+4, season, political
party(Korea, USA), time?), and the number of observations(cases) is 403-
one for each month from 1981 to 2014.

\subparagraph{Brief Description of Variables}
\begin{itemize}
    \item {\tt exchange rate} The market value of the foreign currency reserves held by South Korea. Units: USD Million Current Prices, NSA (monthly) This is our dependent variable.
    \item {\tt gdp} The gross domestic product
    \item {\tt unemployment rate} The percentage of the labor force of South Korea who are unemployed and actively seeking work. Units: Percent (monthly)
    \item {\tt exports} The total value of the goods and services produced by South Korea and purchased by foreign entities. Units: USD Million (Monthly)
    \item {\tt imports} The total value of a country's imports of physical goods and payments to foreigners for services like shipping and tourism. Units: USD Million (Monthly)
    \item {\tt interest rate} The monthly average of the central bank policy rate in South Korea. This is the interest rate the central bank charges on loans to commercial banks. Units: Percent (Monthly)
    \item {\tt inflation rate} The growth rate of the prices in South Korea. (Monthly)
    \item {\tt consumer produce index} The Consumer Price Index (CPI) is a measure of inflation related to the cost of living. It tracks the prices of essential goods and services like food and healthcare, and weighs those price changes according to their relative importance. The overall change in prices is the inflation rate, which can be positive or negative. The standard CPI fails to account for price drops resulting from improved production technologies, so the chained CPI was crafted to allow for a continuously changing basket of goods. Units: Index Points 2010=100, NSA (Monthly)
    \item {\tt debt} The total amount of public and private debt in South Korea owned by foreign creditors. Units: USD Million Current Prices, NSA (quarterly)
    \item {\tt gdp deflator} The relative difference between the real and nominal GDPs of South Korea. Units: Index Points NSA
    \item {\tt goverment spending} The yearly expenditure of the federal government of South Korea. Units: KRW Billion Constant Prices, SA 
    \item {\tt season} categorical
    \item {\tt Political party}  categorical
\end{itemize}

\paragraph{3. Questions of interest}
\begin{itemize}
    \item Can we predict the exchange rate between two countries using various
economic indicators? 
    \item When is the best time to exchange the currency from won(Korea currency) to dollar or vise versa? 
    \item Can we find a predict/relationship equation that could be applied over other countries' currency?

\end{itemize}

\paragraph{4. Plan for the data analysis}
As usual, we start plotting, pair scattor plots, and comparing correlations to pick model candidates. Can we use cross-validation? Since we will have factors(season, predential political parties), could we use 2-way anova the currency rate would different across season, politics?
After obtaining a final model, we also will check if similar models can be
used between other pairs of contries.

\textbf{Anticipated problems}
Some variables like GDP are reported annually or quarterly. To use them with the
variables that are at a monthly grain we will have to impute the missing
values. Some variables in the data are paired (i.e. gdp for Korea, gdp for USA
etc.), so they could be highly correlated. How well do we handle the
correlations? Do we add time as variable?

\end{document}

### 2014-12-03 proposal.tex
starting on missing values

@@
 We use data from Quandl.com, a service that provides clean, documented
-data from a wide variety of sources. The data is loaded directly using a
+data from a wide variety of sources including the US Federal Reserve,
+World Bank, and the National Bank of Korea. The data is loaded directly using a
 web API.
@@
 
-The number of variable set in our data is 22 (+4, season, political
-party(Korea, USA), time?), and the number of observations(cases) is 403-
-one for each month from 1981 to 2014.
+There are 25 variables to analyze. 11 economic indicators for each country,
+as well as season (winter, spring, summer, fall),  and possibly time.
+The number of observations(cases) is 403, one for each month from 1981 to 2014.
+
+{\tt Exchange rate} between Korean Won and US dollar is the primary $Y$
+variable.
 
 \subparagraph{Brief Description of Variables}
+For each country we will examine the following variables.
+
 \begin{itemize}
-    \item {\tt exchange rate} The market value of the foreign currency reserves held by South Korea. Units: USD Million Current Prices, NSA (monthly) This is our dependent variable.
-    \item {\tt gdp} The gross domestic product
-    \item {\tt unemployment rate} The percentage of the labor force of South Korea who are unemployed and actively seeking work. Units: Percent (monthly)
-    \item {\tt exports} The total value of the goods and services produced by South Korea and purchased by foreign entities. Units: USD Million (Monthly)
+    \item {\tt gdp} The gross domestic product. Units: USD Million
+    \item {\tt unemployment rate} The percentage of the labor force who are unemployed and actively seeking work. Units: Percent (monthly)
+    \item {\tt exports} The total value of the goods and services
+produced domestically and purchased by foreign entities. Units: USD Million (Monthly)
     \item {\tt imports} The total value of a country's imports of physical goods and payments to foreigners for services like shipping and tourism. Units: USD Million (Monthly)
-    \item {\tt interest rate} The monthly average of the central bank policy rate in South Korea.
-This is the interest rate the central bank charges on loans to commercial banks.
-Units: Percent (Monthly)
-    \item {\tt inflation rate} The growth rate of the prices in South Korea. (Monthly)
+    \item {\tt interest rate} The monthly average of the central bank policy rate. This is the interest rate the central bank charges on loans to commercial banks. Units: Percent (Monthly)
+    \item {\tt inflation rate} The growth rate of the prices. (Monthly)
     \item {\tt consumer produce index} The Consumer Price Index (CPI) is a measure of inflation related to the cost of living.
@@
 Units: Index Points 2010=100, NSA (Monthly)
-    \item {\tt debt} The total amount of public and private debt in South Korea owned by foreign creditors. Units: USD Million Current Prices, NSA (quarterly)
-    \item {\tt gdp deflator} The relative difference between the real and nominal GDPs of South Korea. Units: Index Points NSA
-    \item {\tt goverment spending} The yearly expenditure of the federal government of South Korea. Units: KRW Billion Constant Prices, SA 
-    \item {\tt season} categorical
-    \item {\tt Political party}  categorical
+    \item {\tt debt} The total amount of public and private debt owned by foreign creditors. Units: USD Million Current Prices, NSA (quarterly)
+    \item {\tt gdp deflator} The relative difference between the real and nominal GDPs. Units: Index Points NSA
+    \item {\tt goverment spending} The yearly expenditure of the federal government. Units: KRW Billion Constant Prices, SA 
+    \item {\tt political party} Categorical variable for the political
+party of the president.
 \end{itemize}

### 2014-12-03 korea.R

@@ -6,9 +6,7 @@ template = read.csv('template.csv', stringsAsFactors=FALSE)
 
 # Target Y variable:
 # Exchange rate: won / USD
-#exchange.rate = Quandl("FRED/EXKOUS", trim_start="1981-04-01", trim_end="2014-10-01", authcode = authcode)
-
-countrycode = 'KOR'
+#exchange = Quandl("FRED/EXKOUS", trim_start="1981-04-01", trim_end="2014-10-01", authcode = authcode)
 
 getcountry = function(countrycode){
     codes = sprintf(template$code, countrycode)
@@ -20,6 +18,33 @@ getcountry = function(countrycode){
     return(out)
 } 
 
-korea = getcountry('KOR')
+#korea = getcountry('KOR')
 
-write.csv(korea, 'korea.csv')
+# Impute missing data
+
+# When does each column start?
+beginning = sapply(korea, function(x) which(!is.na(x))[1])
+
+# So unemployment, interest rate, and debt start later, in 1999
+# That's not so bad. We can truncate the data set to start there and still
+# have about 200 rows. That may actually be better because we cut out the
+# IMF period in 1999.
+
+# Truncated
+k2 = korea[max(beginning):nrow(korea), ]
+
+# May want to chop off the recent end as well
+
+# On which columns do we need to impute?
+needimpute = sapply(k2, function(x) any(is.na(x)))
+
+# Need to plot GDP, gdp deflator, debt, govspending
+# If plots are roughly linear, then imputing through linear
+# interpolation should be ok
+
+# Plot the scaled variables to check for linearity
+sk = scale(k2[, which(needimpute)])
+sk[, 'x'] = 1:nrow(sk)
+
+
+#write.csv(korea, 'korea.csv')

### 2014-12-03 korea.R
linear interpolation of missing data

@@ -1,4 +1,6 @@
 library(Quandl)
+library(reshape2)
+library(ggplot2)
 
 authcode = scan('authcode.txt', what='')
 
@@ -22,8 +24,9 @@ getcountry = function(countrycode){
 
 # Impute missing data
 
-# When does each column start?
-beginning = sapply(korea, function(x) which(!is.na(x))[1])
+# When does each column start and end?
+begindate = sapply(korea, function(x) which(!is.na(x))[1])
+enddate = sapply(korea, function(x) tail(which(!is.na(x)), 1))
 
 # So unemployment, interest rate, and debt start later, in 1999
 # That's not so bad. We can truncate the data set to start there and still
@@ -31,20 +34,39 @@ beginning = sapply(korea, function(x) which(!is.na(x))[1])
 # IMF period in 1999.
 
 # Truncated
-k2 = korea[max(beginning):nrow(korea), ]
-
-# May want to chop off the recent end as well
+k2 = korea[max(begindate):min(enddate), ]
 
 # On which columns do we need to impute?
 needimpute = sapply(k2, function(x) any(is.na(x)))
-
+needimpute[needimpute]
 # Need to plot GDP, gdp deflator, debt, govspending
+which(is.na(k2$interest_rate_KOR))
+# Only one observation missing from interest rate
+
 # If plots are roughly linear, then imputing through linear
-# interpolation should be ok
+# interpolation should be ok. Also could use 
+# exponentially weighted moving average
 
 # Plot the scaled variables to check for linearity
-sk = scale(k2[, which(needimpute)])
-sk[, 'x'] = 1:nrow(sk)
+sk = data.frame(Date = k2$Date, scale(k2[, which(needimpute)]))
+sk_long = reshape2::melt(sk, id = 'Date')
 
+ggplot(data=sk_long[complete.cases(sk_long), ]
+       , aes(x=Date, y=value, color=variable)) + 
+       geom_line()
 
-#write.csv(korea, 'korea.csv')
+# Interest rate stays mostly flat
+# We could add a column that records the number of NA's
+# in that particular row.
+
+num_nas = apply(sk, 1, function(x) sum(is.na(x)))
+plot(as.factor(num_nas))
+# We see that there are lots of NA's. Could add a boolean variable with
+# TRUE for 3 or 4 NA's. May make a difference in regression.
+
+# Now for the linear interpolation. 
+# The Date column was causing problems. Maybe that should be
[…]

### 2014-12-04 korea.R
working with zoo for timeseries

@@ -55,6 +55,8 @@ ggplot(data=sk_long[complete.cases(sk_long), ]
        , aes(x=Date, y=value, color=variable)) + 
        geom_line()
 
+ggsave('figure/scaled_variable_kor.pdf')
+
 # Interest rate stays mostly flat
 # We could add a column that records the number of NA's
 # in that particular row.
@@ -67,6 +69,7 @@ plot(as.factor(num_nas))
 # Now for the linear interpolation. 
 # The Date column was causing problems. Maybe that should be
 # the index?
-kfull = data.frame(na.approx(k2[, -1]))
+kfull = zoo(na.approx(k2[, -1]))
+index(kfull) = k2$Date
 
-write.csv(kfull, 'korea.csv')
+write.zoo(kfull, 'korea.csv')

### 2014-12-05 korea.R
dropped zoo, started using magrittr

@@ -1,6 +1,7 @@
 library(Quandl)
 library(reshape2)
 library(ggplot2)
+library(magrittr)
 
 authcode = scan('authcode.txt', what='')
 
@@ -22,54 +23,52 @@ getcountry = function(countrycode){
 
 #korea = getcountry('KOR')
 
-# Impute missing data
+na_truncate = function(dframe){
+    # Truncates a data frame by identifying the smallest range for
+    # all variables
 
-# When does each column start and end?
-begindate = sapply(korea, function(x) which(!is.na(x))[1])
-enddate = sapply(korea, function(x) tail(which(!is.na(x)), 1))
+    # When does each column start and end?
+    begindate = sapply(dframe, function(x) which(!is.na(x))[1])
+    enddate = sapply(dframe, function(x) tail(which(!is.na(x)), 1))
 
-# So unemployment, interest rate, and debt start later, in 1999
-# That's not so bad. We can truncate the data set to start there and still
-# have about 200 rows. That may actually be better because we cut out the
-# IMF period in 1999.
+    dframe[max(begindate):min(enddate), ]
+}
 
-# Truncated
-k2 = korea[max(begindate):min(enddate), ]
+na_plot = function(dframe, filename){
+    # Plot the scaled variables to check for linearity.
+    # If plots are roughly linear, then imputing through linear
+    # interpolation should be ok.
 
-# On which columns do we need to impute?
-needimpute = sapply(k2, function(x) any(is.na(x)))
-needimpute[needimpute]
-# Need to plot GDP, gdp deflator, debt, govspending
-which(is.na(k2$interest_rate_KOR))
-# Only one observation missing from interest rate
+    # On which columns do we need to impute?
+    needimpute = sapply(dframe, function(x) any(is.na(x)))
 
-# If plots are roughly linear, then imputing through linear
-# interpolation should be ok. Also could use 
-# exponentially weighted moving average
+    scaled = data.frame(Date = dframe$Date, scale(dframe[, which(needimpute)]))
+    scaled_long = reshape2::melt(scaled, id = 'Date')
 
-# Plot the scaled variables to check for linearity
-sk = data.frame(Date = k2$Date, scale(k2[, which(needimpute)]))
-sk_long = reshape2::melt(sk, id = 'Date')
+    ggplot(data=scaled_long[complete.cases(scaled_long), ]
+           , aes(x=Date, y=value, color=variable)) + 
[…]

### 2014-12-05 korea.R
pipeline looks good, now including Quandl step

@@ -12,6 +12,8 @@ template = read.csv('template.csv', stringsAsFactors=FALSE)
 #exchange = Quandl("FRED/EXKOUS", trim_start="1981-04-01", trim_end="2014-10-01", authcode = authcode)
 
 getcountry = function(countrycode){
+    # Fetches live data from Quandl
+
     codes = sprintf(template$code, countrycode)
 
     out = Quandl(codes, trim_start="1981-04-01", trim_end="2014-10-01"
@@ -39,10 +41,13 @@ na_plot = function(dframe, filename){
     # If plots are roughly linear, then imputing through linear
     # interpolation should be ok.
 
-    # On which columns do we need to impute?
-    needimpute = sapply(dframe, function(x) any(is.na(x)))
+    # We'll only impute on the results of na_truncate
+    truncd = na_truncate(dframe)
 
-    scaled = data.frame(Date = dframe$Date, scale(dframe[, which(needimpute)]))
+    # On which columns do we need to impute?
+    needimpute = sapply(truncd, function(x) any(is.na(x)))
+
+    scaled = data.frame(Date = truncd$Date, scale(truncd[, which(needimpute)]))
     scaled_long = reshape2::melt(scaled, id = 'Date')
 
     ggplot(data=scaled_long[complete.cases(scaled_long), ]
@@ -61,13 +66,19 @@ interpolate = function(dframe){
 
 pipecountry = function(country){
     # Takes a single country through the data processing pipeline
+    isocode = 'KOR'
+
+    na_plot(country, sprintf('figure/na_plot_%s.pdf', isocode))
+
     country %>% interpolate %>% na_truncate
 }
 
-a = pipecountry(korea)
+#KOR = pipecountry(korea)
 
 # Now they should be all populated
-sum(complete.cases(a)) == nrow(a)
+sum(complete.cases(KOR)) == nrow(KOR)
+
+save(korea, KOR, file='KOR.rda')
 
 #kfull = data.frame(Date = k2$Date, (na.approx(k2[, -1])))
 

### 2014-12-05 countries.py
Adding main data file country.Rda

@@ -1,0 +1,37 @@
+from requests import get
+import Quandl
+
+baseurl = 'http://api.worldbank.org/countries/all/indicators/'
+
+indicators = ['SP.DYN.SMAM.FE']
+
+fullurl = ''.join((baseurl, indicators[0]))
+
+payload = {'format': 'json', 'date': '2007:2012', 'per_page': '5000'}
+
+response = get(fullurl, params=payload)
+
+j = response.json()[1]
+
+a = []
+for country in j:
+    if country['value']:
+        a.append(country)
+
+def country(name, r=a):
+    '''
+    Pick out one country
+    '''
+    out = []
+    for country in r:
+        if country['country']['value'] == name:
+            out.append(country)
+    return out
+
+
+us = country('United States')
+
+# Try out the Quandl solution
+# Bummed they won't return what I want- a list of this indicator 
+# for all countries!
+#germany = Quandl.get("WORLDBANK/DEU_SP_DYN_SMAM_FE")

### 2014-12-05 country.R

@@ -1,29 +1,49 @@
+# Fri Dec  5 13:48:09 PST 2014
+# Clark Fitzgerald
+# 
+# Collects data from Quandl.com for regression analysis of exchange rates
+# between South Korea and USA.
+# 
+# - Compare other countries by changing the main code in the bottom and
+#   using other 3 letter ISO codes
+#
+# - Save your Quandl authcode in the same directory in a plain text file
+#   called 'authcode.txt'.
+#
+# - Must have a folder in this directory called 'figure' to save plots.
+#
+# - Downloaded and massaged data structures will be saved in 'country.Rda'.
+#   To reload them without downloading again run this in R:
+#       > load('country.Rda')
+#
+# - Edit template.csv to add more parameters using 3 letter ISO codes
+
+
 library(Quandl)
 library(reshape2)
 library(ggplot2)
 library(magrittr)
 
+
+# Date in Quandl format
+today = format(Sys.time(), '%Y-%m-%d')
+
+# Could also put your authcode here as a string
 authcode = scan('authcode.txt', what='')
 
 template = read.csv('template.csv', stringsAsFactors=FALSE)
-
-# Target Y variable:
-# Exchange rate: won / USD
-#exchange = Quandl("FRED/EXKOUS", trim_start="1981-04-01", trim_end="2014-10-01", authcode = authcode)
 
 getcountry = function(countrycode){
     # Fetches live data from Quandl
 
     codes = sprintf(template$code, countrycode)
 
-    out = Quandl(codes, trim_start="1981-04-01", trim_end="2014-10-01"
+    out = Quandl(codes, trim_start="1981-04-01", trim_end=today
                 , collapse='monthly', sort='asc', authcode=authcode)
 
     names(out) = c('Date', paste(template$name, countrycode, sep = '_'))
     return(out)
 } 
-
-#korea = getcountry('KOR')
 
 na_truncate = function(dframe){
     # Truncates a data frame by identifying the smallest range for
@@ -64,22 +84,30 @@ interpolate = function(dframe){
     data.frame(Date = dframe$Date, (na.approx(dframe[, -1])))
 }
 
-pipecountry = function(country){
+pipecountry = function(isocode){
     # Takes a single country through the data processing pipeline
-    isocode = 'KOR'
 
+    country = getcountry(isocode)
     na_plot(country, sprintf('figure/na_plot_%s.pdf', isocode))
 
     country %>% interpolate %>% na_truncate
 }
 
[…]

### 2014-12-05 exploratory.R
exploratory data analysis script

@@ -1,0 +1,4 @@
+# Exploratory data analysis
+
+# Why is debt going down for the US? Because it's a huge negative number!
+# Plot it and look at the range in order to see this.

### 2014-12-06 paper.tex
split data into train and test sets

% A simple template for LaTeX documents
% 
% To produce pdf run:
%   $ pdflatex paper.tex 
%

\documentclass[12pt]{article}

% Begin paragraphs with new line
\usepackage{parskip}

% Change margin size
\usepackage[margin=1in]{geometry}

% Graphics Example:  (PDF's make for good plots)
\usepackage{graphicx}               
% \centerline{\includegraphics{figure.pdf}}

% Allows hyperlinks
\usepackage{hyperref}

% Blocks of code
\usepackage{listings}
\lstset{basicstyle=\ttfamily, title=\lstname}
% Insert code like this. replace `plot.R` with file name.
% \lstinputlisting{plot.R}

% Supports proof environment
\usepackage{amsthm}

% Allows writing \implies and align*
\usepackage{amsmath}

% Allows mathbb{R}
\usepackage{amsfonts}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{center}
    {\Large Assignment 1}\\
    \bigskip
    \hrule
    \medskip
    Clark Fitzgerald\\
    Stats 123
\end{center}

\section*{Makefiles}

Originally designed for compiling large software projects, Makefiles can automate the dependencies between data analysis steps, enabling \textbf{reproducible research}.\footnote{For more on reproducible research, check out the chapter on Open Source Scientific Practice by K. Jarrod Millman and Fernando Perez available at \url{https://osf.io/h9gsd/}.}

\begin{description}
    \item[Statistics-] So hot right now.
    \begin{enumerate}
        \item This is a nested list.
        \item We are shooting for highly nested stuff here.
    \end{enumerate}
    \item[Mathematics-] Always classic.
\end{description}

\vspace{0.3in}

\subsubsection*{R graphics}

Display equations:

\[
    f(x) = x^2 + 5x + \pi
\]

Inline equation $\sum x$, and so on.

You can generate simulated data with R or Python.

\section{R for Graphics}

Here is some R code.

Which generates the corresponding plot:

\centerline{\includegraphics{correlation.pdf}}

\end{document}

### 2014-12-06 correlation.R

@@ -1,0 +1,20 @@
+# We expect that the pairwise correlations will be pretty high
+
+pdf('correlation.pdf')
+load('country.Rda')
+
+corhist = function(dframe){
+    # Plot a histogram of correlations in a dataframe
+
+    # Only doing the numeric columns
+    numcols = sapply(dframe, is.numeric)
+    cormatrix = cor(dframe[, numcols])
+
+    hist(cormatrix[upper.tri(cormatrix)], 
+         main = 'Histogram of Pairwise Correlations',
+         xlab = 'Correlations')
+}
+
+corhist(trainset)
+
+dev.off()

### 2014-12-06 countries.py

@@ -1,37 +1,0 @@
-from requests import get
-import Quandl
-
-baseurl = 'http://api.worldbank.org/countries/all/indicators/'
-
-indicators = ['SP.DYN.SMAM.FE']
-
-fullurl = ''.join((baseurl, indicators[0]))
-
-payload = {'format': 'json', 'date': '2007:2012', 'per_page': '5000'}
-
-response = get(fullurl, params=payload)
-
-j = response.json()[1]
-
-a = []
-for country in j:
-    if country['value']:
-        a.append(country)
-
-def country(name, r=a):
-    '''
-    Pick out one country
-    '''
-    out = []
-    for country in r:
-        if country['country']['value'] == name:
-            out.append(country)
-    return out
-
-
-us = country('United States')
-
-# Try out the Quandl solution
-# Bummed they won't return what I want- a list of this indicator 
-# for all countries!
-#germany = Quandl.get("WORLDBANK/DEU_SP_DYN_SMAM_FE")

### 2014-12-06 download.R

@@ -4,6 +4,9 @@
 # Collects data from Quandl.com for regression analysis of exchange rates
 # between South Korea and USA.
 # 
+# - Downloaded and massaged data structures will be saved in 'country.Rda'.
+#   To reload them without downloading again run this in R:
+#       > load('country.Rda')
 # - Compare other countries by changing the main code in the bottom and
 #   using other 3 letter ISO codes
 #
@@ -12,11 +15,8 @@
 #
 # - Must have a folder in this directory called 'figure' to save plots.
 #
-# - Downloaded and massaged data structures will be saved in 'country.Rda'.
-#   To reload them without downloading again run this in R:
-#       > load('country.Rda')
-#
 # - Edit template.csv to add more parameters using 3 letter ISO codes
+#
 
 
 library(Quandl)
@@ -93,21 +93,46 @@ pipecountry = function(isocode){
     country %>% interpolate %>% na_truncate
 }
 
+ttsplit = function(dframe, testportion, makeglobal=FALSE){
+    # Splits the rows of dframe, returning a list of train and 
+    # validate dataframes 
+    # If makeglobal is TRUE then variables train and validate are created
+    # in the global namespace and nothing is returned.
+
+    n = nrow(country)
+    testsize = round(n * testportion)
+    testindex = sample(1:n, size = testsize, replace=FALSE)
+    out = list(train = country[-testindex, ], 
+               validate = country[testindex, ])
+    if (makeglobal){
+        trainset <<- out[['train']]
+        validate <<- out[['validate']]
+    }
+    else{
+        return(out)
+    }
+}
+
 ############################################################
 # All the action code is here:
 ############################################################
 
-# Target Y variable:
-# Exchange rate: won / USD
-exchange = Quandl("FRED/EXKOUS", trim_start="1981-04-01", trim_end=today
-                  , collapse='monthly', sort='asc', authcode=authcode)
+main = function(){
+    # Target Y variable:
+    # Exchange rate: won / USD
+    exchange = Quandl("FRED/EXKOUS", trim_start="1981-04-01", trim_end=today
+                      , collapse='monthly', sort='asc', authcode=authcode)
 
-names(exchange)[2] = 'exchange_rate'
+    names(exchange)[2] = 'exchange_rate'
 
-USA = pipecountry('USA')
-KOR = pipecountry('KOR')
+    USA = pipecountry('USA')
+    KOR = pipecountry('KOR')
 
[…]

### 2014-12-06 us_debt.R
us debt plot

@@ -1,0 +1,9 @@
+# Why is the US debt decreasing? 
+# Because of the scale - it's quite negative.
+
+pdf('us_debt.pdf')
+load('country.Rda')
+
+plot(USA$Date, USA$debt_USA)
+
+dev.off()

### 2014-12-07 paper.tex
us debt plot
following paper structure

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{document}
 
-\begin{center}
-    {\Large Assignment 1}\\
-    \bigskip
-    \hrule
-    \medskip
-    Clark Fitzgerald\\
-    Stats 123
-\end{center}
+\title{US Korea Exchange Rates \\ STA206 Final Project}
+\date{December 15, 2014}
+\author{Clark Fitzgerald clarkfitzg@gmail.com \\ 
+        Amy Kim atykim@ucdavis.edu}
 
-\section*{Makefiles}
+\maketitle
 
-Originally designed for compiling large software projects, Makefiles can automate the dependencies between data analysis steps, enabling \textbf{reproducible research}.\footnote{For more on reproducible research, check out the chapter on Open Source Scientific Practice by K.
-Jarrod Millman and Fernando Perez available at \url{https://osf.io/h9gsd/}.}
+\begin{abstract}
+    We analyze the exchange rate between the United States and South Korea
+    using monthly country level economic data.
+\end{abstract}
 
-\begin{description}
-    \item[Statistics-] So hot right now.
-    \begin{enumerate}
-        \item This is a nested list.
-        \item We are shooting for highly nested stuff here.
-    \end{enumerate}
-    \item[Mathematics-] Always classic.
-\end{description}
+\section{Introduction}
 
-\vspace{0.3in}
+\section{Methods and Results}
 
-\subsubsection*{R graphics}
+Figure~\ref{fig:correlation} shows a histogram of all the pairwise sample
+correlations.
 
-Display equations:
+Figure~\ref{fig:na_plot_KOR} shows the variables which were interpolated.
 
-\[
-    f(x) = x^2 + 5x + \pi
-\]
+\section{Conclusions and Discussion}
 
-Inline equation $\sum x$, and so on.
+\newpage
+\section{Appendix}
 
-You can generate simulated data with R or Python.
+\listoffigures
 
-\section{R for Graphics}
+\begin{figure}
+  \centering
+    \includegraphics{correlation.pdf}
+  \caption{Histogram of all pairwise sample correlations.}
+  \label{fig:correlation}
+\end{figure}
 
-Here is some R code.
-
-Which generates the corresponding plot:
-
-\centerline{\includegraphics{correlation.pdf}}
+\begin{figure}
+  \centering
+    \includegraphics{na_plot_KOR.pdf}
+  \caption{Scaled variables with NA values.}
+  \label{fig:na_plot_KOR}
+\end{figure}
 

### 2014-12-07 na_plot_KOR.R

@@ -1,0 +1,18 @@
+# Justification of linear interpolation
+library(ggplot2)
+library(reshape2)
+
+load('country.Rda')
+
+imputed = c('gdp_KOR', 'gdp_deflator_KOR', 'debt_KOR', 
+            'interest_rate_KOR', 'govspending_KOR')
+
+scaled = data.frame(Date = KOR$Date, scale(KOR[, imputed]))
+
+scaled_long = reshape2::melt(scaled, id = 'Date')
+
+ggplot(data=scaled_long[complete.cases(scaled_long), ]
+       , aes(x=Date, y=value, color=variable)) + 
+       geom_line()
+
+ggsave('na_plot_KOR.pdf', height = 4, units = 'in')

### 2014-12-07 correlation.R
remove title from correlation plot

@@ -10,8 +10,7 @@ corhist = function(dframe){
     numcols = sapply(dframe, is.numeric)
     cormatrix = cor(dframe[, numcols])
 
-    hist(cormatrix[upper.tri(cormatrix)], 
-         main = 'Histogram of Pairwise Correlations',
+    hist(cormatrix[upper.tri(cormatrix)], main = '',
          xlab = 'Correlations')
 }
 

### 2014-12-08 paper.tex
adding reference page
introductory and background material

@@
     We analyze the exchange rate between the United States and South Korea
-    using monthly country level economic data.
+    (hereafter Korea) using monthly country level economic data.
 \end{abstract}
@@
 
+\subsection{Background}
+
+In today's highly connected global society people and money move between
+countries. The exchange rates between countries determine the value of
+that currency.
+
+We analyze data on exchange rate between the US and Korea from 1999 to
+2014. In 1997 the IMF crisis in Korea perturbed the economic indicators,
+but in looking at the data we can see that it has been stable since 1999.
+
+\subsection{Questions}
+
+\begin{enumerate}
+
+    \item How do exchange rates behave?
+
+    \item Can we predict the exchange rate between two countries?
+
+    \item When is the best time to exchange currency from won (Korean
+        currency) to dollar or vice versa?
+
+    \item Can we find a linear relationship
+        that can be applied over other countries' currency?
+
+\end{enumerate}
+
+\subsection{Motivation}
+
+Sometimes it's necessary to transfer funds between countries. This is true
+for both individuals, families, and organizations. The exchange rate can
+vary substantially over time. For example, in
+Figure~\ref{fig:exchange_rate} we observe the exchange rate between Korea
+and the US varying between 900 and 1400 Korean Won per US Dollar. A quick
+calculation on this data 
+implies that during this period a fixed amount of Korean currency 
+could have been worth \$100,000 or \$158,000, depending on when it was  
+exchanged.
+
+Hence if one has a significant amount of capital to move from
+one country to another and some flexibility around the timing then it makes
+sense to do it when the exchange rates
+are favorable for the transfer. For example, you may want to sell a
+house in Korea and put that money towards a house in the US.
+
+If one plans to go on vacation in a foreign country then it makes sense to
+visit a country where the exchange rate is favorable.
+
+\subsection{Data}
+We use data from Quandl.com, a service that provides clean, documented
+data from a wide variety of sources including the US Federal Reserve,
+World Bank, and the National Bank of Korea. 
+{\tt Exchange rate} between Korean Won and US dollar is the primary $Y$
+variable.
+
+For each country we will examine the following variables:
+
+\begin{itemize}
+    \item {\tt gdp} The gross domestic product. Units: USD Million
+    \item {\tt unemployment rate} The percentage of the labor force who are unemployed and actively seeking work. Units: Percent (monthly)
+    \item {\tt exports} The total value of the goods and services
+produced domestically and purchased by foreign entities. Units: USD Million (Monthly)
+    \item {\tt imports} The total value of a country's imports of physical goods and payments to foreigners for services like shipping and tourism. Units: USD Million (Monthly)
+    \item {\tt interest rate} The monthly average of the central bank policy rate. This is the interest rate the central bank charges on loans to commercial banks. Units: Percent (Monthly)
+    \item {\tt inflation rate} The growth rate of the prices. (Monthly)
+    \item {\tt consumer produce index} The Consumer Price Index (CPI) is a measure of inflation related to the cost of living. Units: Index Points 2010=100, NSA (Monthly)
+    \item {\tt debt} The total amount of public and private debt owned by foreign creditors. Units: USD Million Current Prices, NSA (quarterly)
+    \item {\tt gdp deflator} The relative difference between the real and nominal GDPs. Units: Index Points NSA
+    \item {\tt goverment spending} The yearly expenditure of the federal
+        government. Units: local currency
+    \item {\tt political party} Categorical variable for the political
+party of the president.
+\end{itemize}
+
 \section{Methods and Results}
+
+\subsection{Data Collection}
+The data is loaded directly using a
+web API (Application programming interface).
+Using this service will make it easier to repeat the analysis in the future, or conduct similar analyses between
+different countries.
 
@@
 Figure~\ref{fig:na_plot_KOR} shows the variables which were interpolated.
+
+Attempting to reference \cite{lamport94}.
 
@@
   \centering
+    \includegraphics{exchange_rate.pdf}
+  \caption{Exchange rate between US and South Korea from 1999 to 2014}
+  \label{fig:exchange_rate}
+\end{figure}
+
+\begin{figure}
+  \centering
     \includegraphics{correlation.pdf}
-  \caption{Histogram of all pairwise sample correlations.}
+  \caption{Histogram of all pairwise sample correlations}
   \label{fig:correlation}
@@
     \includegraphics{na_plot_KOR.pdf}
-  \caption{Scaled variables with NA values.}
+  \caption{Scaled variables with NA values}
   \label{fig:na_plot_KOR}
@@
 
+\newpage
+
+\begin{thebibliography}{9}
+
+\bibitem{lamport94}
+  Leslie Lamport,
+  \emph{\LaTeX: a document preparation system}.
+  Addison Wesley, Massachusetts,
+  2nd edition,
+  1994.
+
+\end{thebibliography}
+
 \end{document}

### 2014-12-08 exploratory.R

@@ -1,4 +1,15 @@
+library(ggplot2)
+
 # Exploratory data analysis
+load('country.Rda')
 
 # Why is debt going down for the US? Because it's a huge negative number!
 # Plot it and look at the range in order to see this.
+
+# Exchange rate over time
+qplot(Date, exchange_rate, data = country, ylab = 'Won / Dollar')
+ggsave('exchange_rate.pdf', width = 6, height = 8, units = 'in')
+
+# What's the max difference?
+r = range(country$exchange_rate)
+r[2] / r[1]

### 2014-12-09 download.R
implemented local caching, decoupled download step

@@ -15,15 +15,11 @@
 #
 # - Must have a folder in this directory called 'figure' to save plots.
 #
-# - Edit template.csv to add more parameters using 3 letter ISO codes
+# - Edit quandl_template.csv to add more parameters using 3 letter ISO codes
 #
 
 
 library(Quandl)
-library(reshape2)
-library(ggplot2)
-library(magrittr)
-
 
 # Date in Quandl format
 today = format(Sys.time(), '%Y-%m-%d')
@@ -31,7 +27,7 @@ today = format(Sys.time(), '%Y-%m-%d')
 # Could also put your authcode here as a string
 authcode = scan('authcode.txt', what='')
 
-template = read.csv('template.csv', stringsAsFactors=FALSE)
+template = read.csv('quandl_template.csv', stringsAsFactors=FALSE)
 
 getcountry = function(countrycode){
     # Fetches live data from Quandl
@@ -45,94 +41,14 @@ getcountry = function(countrycode){
     return(out)
 } 
 
-na_truncate = function(dframe){
-    # Truncates a data frame by identifying the smallest range for
-    # all variables
+# Target Y variable:
+# Exchange rate: won / USD
+exchange = Quandl("FRED/EXKOUS", trim_start="1981-04-01", trim_end=today
+                  , collapse='monthly', sort='asc', authcode=authcode)
 
-    # When does each column start and end?
-    begindate = sapply(dframe, function(x) which(!is.na(x))[1])
-    enddate = sapply(dframe, function(x) tail(which(!is.na(x)), 1))
+names(exchange)[2] = 'exchange_rate'
 
-    dframe[max(begindate):min(enddate), ]
-}
+USA = getcountry('USA')
+KOR = getcountry('KOR')
 
-na_plot = function(dframe, filename){
-    # Plot the scaled variables to check for linearity.
-    # If plots are roughly linear, then imputing through linear
-    # interpolation should be ok.
-
-    # We'll only impute on the results of na_truncate
-    truncd = na_truncate(dframe)
-
-    # On which columns do we need to impute?
-    needimpute = sapply(truncd, function(x) any(is.na(x)))
-
-    scaled = data.frame(Date = truncd$Date, scale(truncd[, which(needimpute)]))
-    scaled_long = reshape2::melt(scaled, id = 'Date')
-
-    ggplot(data=scaled_long[complete.cases(scaled_long), ]
-           , aes(x=Date, y=value, color=variable)) + 
[…]

### 2014-12-09 exchange_rate.R

@@ -1,0 +1,9 @@
+# Exploratory data analysis
+
+load('country.Rda')
+
+pdf('exchange_rate.pdf')
+
+with(country, plot(Date, exchange_rate, ylab = 'Won / Dollar'))
+
+dev.off()

### 2014-12-09 na_plot_KOR.R

@@ -2,6 +2,28 @@
 library(ggplot2)
 library(reshape2)
 
+na_plot = function(dframe, filename){
+    # Plot the scaled variables to check for linearity.
+    # If plots are roughly linear, then imputing through linear
+    # interpolation should be ok.
+
+    # We'll only impute on the results of na_truncate
+    truncd = na_truncate(dframe)
+
+    # On which columns do we need to impute?
+    needimpute = sapply(truncd, function(x) any(is.na(x)))
+
+    scaled = data.frame(Date = truncd$Date, scale(truncd[, which(needimpute)]))
+    scaled_long = reshape2::melt(scaled, id = 'Date')
+
+    ggplot(data=scaled_long[complete.cases(scaled_long), ]
+           , aes(x=Date, y=value, color=variable)) + 
+           geom_line()
+
+    ggsave(filename)
+}
+
+
 load('country.Rda')
 
 imputed = c('gdp_KOR', 'gdp_deflator_KOR', 'debt_KOR', 

### 2014-12-09 preprocess.R

@@ -1,0 +1,71 @@
+# Preprocessing steps
+
+library(reshape2)
+library(ggplot2)
+library(magrittr)
+
+
+na_truncate = function(dframe){
+    # Truncates a data frame by identifying the smallest range for
+    # all variables
+
+    # When does each column start and end?
+    begindate = sapply(dframe, function(x) which(!is.na(x))[1])
+    enddate = sapply(dframe, function(x) tail(which(!is.na(x)), 1))
+
+    dframe[max(begindate):min(enddate), ]
+}
+
+interpolate = function(dframe){
+    # Impute missing data using linear interpolation.
+
+    # The Date column was causing problems.
+    data.frame(Date = dframe$Date, (na.approx(dframe[, -1])))
+}
+
+pipecountry = function(isocode){
+    # Takes a single country through the data processing pipeline
+
+    country = getcountry(isocode)
+    na_plot(country, sprintf('figure/na_plot_%s.pdf', isocode))
+
+    country %>% interpolate %>% na_truncate
+}
+
+ttsplit = function(dframe, testportion, makeglobal=FALSE){
+    # Splits the rows of dframe, returning a list of train and 
+    # validate dataframes 
+    # If makeglobal is TRUE then variables train and validate are created
+    # in the global namespace and nothing is returned.
+
+    n = nrow(country)
+    testsize = round(n * testportion)
+    testindex = sample(1:n, size = testsize, replace=FALSE)
+    out = list(train = country[-testindex, ], 
+               validate = country[testindex, ])
+    if (makeglobal){
+        trainset <<- out[['train']]
+        validate <<- out[['validate']]
+    }
+    else{
+        return(out)
+    }
+}
+
+x = 10
+save(x, file='country.Rda')
+############################################################
+# All the action code is here:
+############################################################
+#
+#USA = pipecountry('USA')
+#KOR = pipecountry('KOR')
+#
+## Merge on common 'Date' column. This is the primary table for analysis
+#country = merge(exchange, merge(KOR, USA))
+#
+## Reserve one third of the data for a validation set
+#ttsplit(country, testportion = 1/3, makeglobal=TRUE)
+#
+#save(validate, trainset, country, exchange, USA, KOR, file='country.Rda')
+#}

### 2014-12-09 exploratory.R
finished preprocessing and started preliminary analysis

@@ -13,3 +13,18 @@ ggsave('exchange_rate.pdf', width = 6, height = 8, units = 'in')
 # What's the max difference?
 r = range(country$exchange_rate)
 r[2] / r[1]
+
+# Will it fit?
+fit1 = lm(exchange_rate ~ ., data = trainset[, -1])
+summary(fit1)
+plot(fit1)
+
+# Intentionally Remove the effects of time
+timefit = lm(exchange_rate ~ Date, data=trainset)
+plot(timefit)
+
+country_notime = cbind(exchange_notime = trainset$exchange_rate - predict(timefit),
+                       trainset[, -c(1, 2)])
+fit2 = lm(exchange_notime ~ ., data = country_notime)
+summary(fit2)
+plot(fit2)

### 2014-12-09 preprocess.R

@@ -1,5 +1,6 @@
 # Preprocessing steps
 
+library(zoo)
 library(reshape2)
 library(ggplot2)
 library(magrittr)
@@ -20,14 +21,12 @@ interpolate = function(dframe){
     # Impute missing data using linear interpolation.
 
     # The Date column was causing problems.
-    data.frame(Date = dframe$Date, (na.approx(dframe[, -1])))
+    data.frame(Date = dframe$Date, (zoo::na.approx(dframe[, -1])))
 }
 
-pipecountry = function(isocode){
+pipecountry = function(country){
     # Takes a single country through the data processing pipeline
-
-    country = getcountry(isocode)
-    na_plot(country, sprintf('figure/na_plot_%s.pdf', isocode))
+    #na_plot(country, sprintf('figure/na_plot_%s.pdf', isocode))
 
     country %>% interpolate %>% na_truncate
 }
@@ -52,20 +51,16 @@ ttsplit = function(dframe, testportion, makeglobal=FALSE){
     }
 }
 
-x = 10
-save(x, file='country.Rda')
-############################################################
-# All the action code is here:
-############################################################
-#
-#USA = pipecountry('USA')
-#KOR = pipecountry('KOR')
-#
-## Merge on common 'Date' column. This is the primary table for analysis
-#country = merge(exchange, merge(KOR, USA))
-#
-## Reserve one third of the data for a validation set
-#ttsplit(country, testportion = 1/3, makeglobal=TRUE)
-#
-#save(validate, trainset, country, exchange, USA, KOR, file='country.Rda')
-#}
+# Cache from downloader
+load('cache.Rda')
+
+USA_clean = pipecountry(USA)
+KOR_clean = pipecountry(KOR)
+
+# Merge on common 'Date' column. This is the primary table for analysis
+country = merge(exchange, merge(KOR_clean, USA_clean))
+
+# Reserve one third of the data for a validation set
+ttsplit(country, testportion = 1/3, makeglobal=TRUE)
+
+save(validate, trainset, country, exchange, USA, KOR, file='country.Rda')

### 2014-12-09 exploratory.R
attempted to remove linear effect of time from response

@@ -22,6 +22,8 @@ plot(fit1)
 # Intentionally Remove the effects of time
 timefit = lm(exchange_rate ~ Date, data=trainset)
 plot(timefit)
+# This is a terrible fit - exchange rate is not linear over time.
+# Anyways, it would be difficult to fully remove time
 
 country_notime = cbind(exchange_notime = trainset$exchange_rate - predict(timefit),
                        trainset[, -c(1, 2)])

### 2014-12-11 paper.tex
reproducibility section

@@
 In today's highly connected global society people and money move between
-countries. The exchange rates between countries determine the value of
-that currency.
+countries. The exchange rates between countries determine the relative
+value of that money.
 
@@
 
-If one plans to go on vacation in a foreign country then it makes sense to
-visit a country where the exchange rate is favorable.
+\subsection{Data}
 
-\subsection{Data}
 We use data from Quandl.com, a service that provides clean, documented
@@
 
-\subsection{Data Collection}
-The data is loaded directly using a
-web API (Application programming interface).
-Using this service will make it easier to repeat the analysis in the future, or conduct similar analyses between
+\subsection{Dynamic Data}
+
+The data used in this analysis is unusual- it's generated dynamically from
+live, high quality sources, and is self updating. The data is loaded directly from Quandl using a REST (Representational 
+State Transfer) web API (Application programming interface), and then
+cached locally, limiting network dependence.
+Using this service makes it easy to repeat the analysis in the future, or conduct similar analyses between
 different countries.
+
+\subsection{Reproducibility}
+
+Above we mentioned that the data and all analysis in the report is self 
+updating. This is accomplished through the use of GNU \texttt{make}, which
+describes a DAG (Directed Acyclic Graph) of dependencies for the final
+report. \texttt{make} detects file modifications and will lazily run all
+commands and scripts required for the final output.
+
+What does this mean? Suppose that we need more variables to do the
+regression. Thanks to the use
+of industry standard ISO codes for the countries and proper parameters in
+our code base this becomes a simple task.
+We've stored string templates for the country level 
+variables in a file
+called \texttt{template.txt}. The data collection and preprocessing steps
+are automated, so we can add more variables simply by adding a row to
+\texttt{template.txt} and entering the command \texttt{make} from the shell
+prompt, which causes the following sequence of events:
+
+\begin{enumerate}
+    \item \texttt{make} detects the changed template file
+    \item The script \texttt{download.R} runs, updating the cache
+    \item The script \texttt{preprocess.R} runs, applying the appropriate
+        transformations and joining the tables to a form suitable for
+        analysis
+    \item All other scripts which depend on the preprocessed data run
+    \item The report output is produced
+\end{enumerate}
+
+In this case the output is a PDF, but it could just as easily be a web page
+on the department server, or an upload into another REST API.
+
+If we were doing this analysis for a client and need to run it again a year
+later with a new year of data 
+then all we have to do is type \texttt{make}. We'll still need to interpret
+plots and models, but the most time consuming work is done.
+
+One other aspect of reproducibility is version control. In this
+project we've used \texttt{Git} to collect snapshots of the project at
+every stage. The project is hosted publicly on
+Github\footnote{\url{https://github.com/clarkfitzg/stats206\_project}}.
+A look at the log file shows what work happened when and why,
+which is important for establishing provenance. In particular it shows 
+that we've been active on this project every day since
+December 1st. Here's an example of a log entry:
+
+\begin{verbatim}
+commit e2beaf01b0a1457e7167989c1fae575c2a676f19
+Author: Clark Fitzgerald <clarkfitzg@gmail.com>
+Date:   Mon Dec 8 20:51:04 2014 -0800
+
+    implemented local caching, decoupled download step
+\end{verbatim}
+The point is that through a little engineering the data analysis process
+can be made transparent, automated, extensible, and fully reproducible.
 

### 2014-12-11 paper.tex
set seed for preprocess test train split

@@
 \end{verbatim}
-The point is that through a little engineering the data analysis process
+
+Through a little engineering the data analysis process
 can be made transparent, automated, extensible, and fully reproducible.
+
+Modern high level programming languages like R provide incredible tools. By
+using their capabilities we can
 

### 2014-12-11 preprocess.R

@@ -6,6 +6,8 @@ library(ggplot2)
 library(magrittr)
 
 
+set.seed(893)
+
 na_truncate = function(dframe){
     # Truncates a data frame by identifying the smallest range for
     # all variables

### 2014-12-11 exploratory.R
adding file for functions

@@ -2,6 +2,7 @@ library(ggplot2)
 
 # Exploratory data analysis
 load('country.Rda')
+source('functions.R')
 
 # Why is debt going down for the US? Because it's a huge negative number!
 # Plot it and look at the range in order to see this.
@@ -27,6 +28,28 @@ plot(timefit)
 
 country_notime = cbind(exchange_notime = trainset$exchange_rate - predict(timefit),
                        trainset[, -c(1, 2)])
+
 fit2 = lm(exchange_notime ~ ., data = country_notime)
 summary(fit2)
 plot(fit2)
+
+
+# Thu Dec 11 14:18:14 PST 2014
+
+set.seed(123)
+
+# Get the difference between each previous row
+countrydiff = as.data.frame(sapply(country, diff))
+
+# Drop the date column
+countrydiff = countrydiff[, -1]
+
+ttsplit(countrydiff, 1 / 3, makeglobal=TRUE)
+
+mod1 = lm(exchange_rate ~ ., data=trainset)
+summary(mod1)
+
+library(leaps)
+
+r1 = regsubsets(exchange_rate ~ ., data=trainset)
+s1 = summary(r1)

### 2014-12-11 functions.R

@@ -1,0 +1,52 @@
+# Library of useful analytic functions
+
+library(zoo)
+library(reshape2)
+library(ggplot2)
+library(magrittr)
+
+
+na_truncate = function(dframe){
+    # Truncates a data frame by identifying the smallest range for
+    # all variables
+
+    # When does each column start and end?
+    begindate = sapply(dframe, function(x) which(!is.na(x))[1])
+    enddate = sapply(dframe, function(x) tail(which(!is.na(x)), 1))
+
+    dframe[max(begindate):min(enddate), ]
+}
+
+interpolate = function(dframe){
+    # Impute missing data using linear interpolation.
+
+    # The Date column was causing problems.
+    data.frame(Date = dframe$Date, (zoo::na.approx(dframe[, -1])))
+}
+
+pipecountry = function(country){
+    # Takes a single country through the data processing pipeline
+    #na_plot(country, sprintf('figure/na_plot_%s.pdf', isocode))
+
+    country %>% interpolate %>% na_truncate
+}
+
+ttsplit = function(dframe, testportion, makeglobal=FALSE){
+    # Splits the rows of dframe, returning a list of train and 
+    # validate dataframes 
+    # If makeglobal is TRUE then variables train and validate are created
+    # in the global namespace and nothing is returned.
+
+    n = nrow(dframe)
+    testsize = round(n * testportion)
+    testindex = sample(1:n, size = testsize, replace=FALSE)
+    out = list(train = dframe[-testindex, ], 
+               validate = dframe[testindex, ])
+    if (makeglobal){
+        trainset <<- out[['train']]
+        validate <<- out[['validate']]
+    }
+    else{
+        return(out)
+    }
+}

### 2014-12-11 preprocess.R

@@ -1,60 +1,12 @@
 # Preprocessing steps
 
-library(zoo)
-library(reshape2)
-library(ggplot2)
-library(magrittr)
+# Cache from downloader
+load('cache.Rda')
+
+source('functions.R')
 
 
 set.seed(893)
-
-na_truncate = function(dframe){
-    # Truncates a data frame by identifying the smallest range for
-    # all variables
-
-    # When does each column start and end?
-    begindate = sapply(dframe, function(x) which(!is.na(x))[1])
-    enddate = sapply(dframe, function(x) tail(which(!is.na(x)), 1))
-
-    dframe[max(begindate):min(enddate), ]
-}
-
-interpolate = function(dframe){
-    # Impute missing data using linear interpolation.
-
-    # The Date column was causing problems.
-    data.frame(Date = dframe$Date, (zoo::na.approx(dframe[, -1])))
-}
-
-pipecountry = function(country){
-    # Takes a single country through the data processing pipeline
-    #na_plot(country, sprintf('figure/na_plot_%s.pdf', isocode))
-
-    country %>% interpolate %>% na_truncate
-}
-
-ttsplit = function(dframe, testportion, makeglobal=FALSE){
-    # Splits the rows of dframe, returning a list of train and 
-    # validate dataframes 
-    # If makeglobal is TRUE then variables train and validate are created
-    # in the global namespace and nothing is returned.
-
-    n = nrow(country)
-    testsize = round(n * testportion)
-    testindex = sample(1:n, size = testsize, replace=FALSE)
-    out = list(train = country[-testindex, ], 
-               validate = country[testindex, ])
-    if (makeglobal){
-        trainset <<- out[['train']]
-        validate <<- out[['validate']]
-    }
-    else{
-        return(out)
-    }
-}
-
-# Cache from downloader
-load('cache.Rda')
 
 USA_clean = pipecountry(USA)
 KOR_clean = pipecountry(KOR)

### 2014-12-11 exploratory.R
Further exploratory work

@@ -38,13 +38,52 @@ plot(fit2)
 
 set.seed(123)
 
-# Get the difference between each previous row
-countrydiff = as.data.frame(sapply(country, diff))
-
 # Drop the date column
-countrydiff = countrydiff[, -1]
+nodate = country[, names(country) != 'Date']
+dim(nodate)
+
+# Exploratory data analysis
+pdf('scatterplot.pdf')
+plot(nodate[, 1:5])
+dev.off()
+
+plot(nodate[, 1:5])
+# Time dependent patterns are very obvious here
+# We want to be seeing a cloud
+# Who put these snakes in my data?!!
+plot(nodate[, c(1, 6:10)])
+plot(nodate[, c(1, 11:16)])
+plot(nodate[, c(1, 17:21)])
+
+# Suppose we pick out the variables where the patterns between
+# exchange_rate and the other variables are least obvious?
+# We can see them from the plots above.
+nopattern = c('exchange_rate', 'unemployment_KOR', 'interest_rate_KOR',
+              'inflation_KOR', 'interest_rate_USA', 'inflation_USA')
+
+# What happens if we just regress on these?
+fitnopattern = lm(exchange_rate ~ ., data = nodate[, nopattern])
+summary(fitnopattern)
+# R^2 is around 0.6, so it doesn't do all that well.
+
+# Get the difference between each previous row
+# As suggested by Taeyen
+countrydiff = as.data.frame(sapply(nodate, diff))
 
 ttsplit(countrydiff, 1 / 3, makeglobal=TRUE)
+
+plot(trainset[, c('exchange_rate', 'interest_rate_KOR', 'interest_rate_USA')]) 
+
+# Many of these 'diffed' variables are close to discrete
+
+plot(trainset[, 1:5])
+# This looks much more like a proper data cloud
+# But I worry that exchange rate is too flat when seen as a function of the
+# other variables
+plot(trainset[, c(1, 6:10)])
+plot(trainset[, c(1, 11:16)])
+plot(trainset[, c(1, 17:21)])
+
 
 mod1 = lm(exchange_rate ~ ., data=trainset)
 summary(mod1)
@@ -53,3 +92,17 @@ library(leaps)
 
 r1 = regsubsets(exchange_rate ~ ., data=trainset)
 s1 = summary(r1)
+
+showmodel = function(p, leapsummary){
+    # returns the coefficient names from the model with p coefficients
+    # p - 1 used because first term includes intercept
[…]

### 2014-12-12 exploratory.R
Model building

@@ -90,8 +90,49 @@ summary(mod1)
 
 library(leaps)
 
-r1 = regsubsets(exchange_rate ~ ., data=trainset)
-s1 = summary(r1)
+rs1 = regsubsets(exchange_rate ~ ., data=trainset, nbest=1, nvmax=21)
+rs1s = summary(rs1)
+
+modelstats = data.frame(p = 2:21,
+    sse = rs1s$rss,
+    r2 = rs1s$rsq,
+    r2a = rs1s$adjr2,
+    cp = rs1s$cp)
+
+modelstats
+
+
+fullmod = mod1
+fullMSE = summary(fullmod)$sigma ** 2
+
+ssto = sum((trainset$exchange_rate - mean(trainset$exchange_rate)) ** 2)
+
+n = nrow(trainset)
+
+############################################################
+# SSE should be decreasing and nonnegative
+# Very surprisingly, rss here is what we have been calling sse!
+modelstats = data.frame(p = 1:ncol(trainset) + 1,
+    sse = rs1s$rss,
+    r2 = rs1s$rsq,
+    r2a = rs1s$adjr2,
+    cp = rs1s$cp)
+
+# define the first row
+row1 = data.frame(p = 1, sse = ssto, r2 = 0, r2a = 0,
+                  cp = ssto / fullMSE - (n - 2))
+
+# Add the first row with the others
+modelstats = rbind(row1, modelstats)
+
+# Add AIC, BIC columns
+modelstats$aic = with(modelstats, n * log(sse / n) + 2 * p)
+modelstats$bic = with(modelstats, n * log(sse / n) + log(n) * p)
+
+modelstats
+sapply(modelstats, which.min)
+#
+
 
 showmodel = function(p, leapsummary){
     # returns the coefficient names from the model with p coefficients
@@ -103,6 +144,128 @@ showmodel = function(p, leapsummary){
 # Only picked 9- and they look like the ones that were more of a cloud from
 # the EDA- the nopattern ones
 nopattern
-sapply(1:9, showmodel, s1)
+sapply(1:9, showmodel, rs1s)
 
+hist(country$inflation_USA)
+
+# Try adding KOR / USD variables
+ratio = with(country, 
+             data.frame(exchange_rate = exchange_rate
+                        , gdp = gdp_KOR / gdp_USA
+                        , unemployment = unemployment_KOR / unemployment_USA
[…]

### 2014-12-12 paper.tex
Further exploratory work
Discovered bug in Quandl data

@@
 
+\subsection{Exploratory Data Analysis}
+
+We started out with some general summary statistics and plotting. The first
+issue was missing data, which we had anticipated. While most of the
+variables were reported monthly, \texttt{debt} was reported on a quarterly
+basis and 
+Our simple approach for dealing with this was to
+make a plot of these variables over time.
+If they are approximately linear then we could 
+linearly interpolate with respect to time. The scaled plots 
+\ref{fig:na_plot_KOR} and
+\ref{fig:na_plot_USA} do appear to be mostly linear, so we interpolated. 
+In this plots one can observe that $debt_USA$ is decreasing; this is
+because $debt$ carries a sign. The value of $debt_USA$ on
+
+When beginning this project we knew that the time dependency and
+multicollinearity would be the
+biggest obstacles, since in this case the assumption that the errors
+$\epsilon_i \sim N(0, \sigma^2)$ does not hold.
+Figure~\ref{fig:scatterplot} shows what the original data looks like.
+We observe clear patterns emerging, with some even approaching a
+continuous mathematical functional relationship- ie the patterns
+between \texttt{gdp}, \texttt{gdp\_deflator}, and \texttt{debt} for Korea.
 Figure~\ref{fig:correlation} shows a histogram of all the pairwise sample
-correlations.
+correlations for the original data set- they are
+
+Our first approach for overcoming this time dependence was to create a new
+data frame from the original data by taking the difference between each
+subsequent row. More formally, if our original design matrix $X$ is $n
+\times p$ then we can define a new $n - 1 \times p$ matrix $\tilde{X}$ by
+setting
+\[
+    \tilde{X_{i,*}} = X_{(i + 1), *} - X_{i, *}
+\]
+for $i = 1, 2, \dots, n-1$ where $X_{i, *}$ represents the $i$th row of
+$X$.
+
+This approach did indeed eliminate the most obvious patterns, but 
+sacrifices interpretability
+
+Our first step was to separate the data into training and validation sets,
+and only
 
@@
   \centering
-    \includegraphics{correlation.pdf}
-  \caption{Histogram of all pairwise sample correlations}
-  \label{fig:correlation}
+    \includegraphics{na_plot_KOR.pdf}
+  \caption{Scaled variables with NA values}
+  \label{fig:na_plot_KOR}
 \end{figure}
@@
   \centering
-    \includegraphics{na_plot_KOR.pdf}
+    \includegraphics{na_plot_USA.pdf}
   \caption{Scaled variables with NA values}
-  \label{fig:na_plot_KOR}
+  \label{fig:na_plot_USA}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics{scatterplot.pdf}
+  \caption{Scatterplot of untransformed data}
+  \label{fig:scatterplot}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics{correlation.pdf}
+  \caption{Histogram of all pairwise sample correlations}
+  \label{fig:correlation}
 \end{figure}

### 2014-12-12 analysis.R

@@ -1,0 +1,39 @@
+# Final analysis to produce report
+
+library(ggplot2)
+
+set.seed(37)
+
+source('functions.R')
+load('country.Rda')
+
+############################################################
+# Exploratory data analysis
+############################################################
+
+# Exchange rate over time
+qplot(Date, exchange_rate, data = country, ylab = 'Won / Dollar')
+ggsave('exchange_rate.pdf', width = 6, height = 8, units = 'in')
+
+# Most recent US debt figures
+# Something is very wrong, starting in 2014-05-31
+# Debt is way too high.
+tail(USA[!is.na(USA$debt_USA), c('Date', 'debt_USA')])
+
+# Justifies linear interpolation
+na_plot(KOR)
+ggsave('na_plot_KOR.pdf', height = 4, units = 'in')
+na_plot(USA)
+ggsave('na_plot_USA.pdf', height = 4, units = 'in')
+
+# Drop the date column, so we know that it's not used in the analysis
+nodate = country[, names(country) != 'Date']
+dim(nodate)
+
+pdf('scatterplot.pdf')
+plot(nodate[, 1:5])
+dev.off()
+
+pdf('correlation.pdf')
+corhist(trainset)
+dev.off()

### 2014-12-12 functions.R

@@ -50,3 +50,35 @@ ttsplit = function(dframe, testportion, makeglobal=FALSE){
         return(out)
     }
 }
+
+corhist = function(dframe){
+    # Plot a histogram of correlations in a dataframe
+
+    # Only doing the numeric columns
+    numcols = sapply(dframe, is.numeric)
+    cormatrix = cor(dframe[, numcols])
+
+    hist(cormatrix[upper.tri(cormatrix)], main = '',
+         xlab = 'Correlations')
+}
+
+na_plot = function(dframe, filename){
+    # Plot the scaled variables to check for linearity.
+    # If plots are roughly linear, then imputing through linear
+    # interpolation should be ok.
+
+    # We'll only impute on the results of na_truncate
+    truncd = na_truncate(dframe)
+
+    # On which columns do we need to impute?
+    needimpute = sapply(truncd, function(x) any(is.na(x)))
+
+    scaled = data.frame(Date = truncd$Date, scale(truncd[, which(needimpute)]))
+    scaled_long = reshape2::melt(scaled, id = 'Date')
+
+    p = ggplot(data=scaled_long[complete.cases(scaled_long), ]
+           , aes(x=Date, y=value, color=variable)) + 
+           geom_line()
+
+    return(p)
+}

### 2014-12-13 paper.tex
report

@@
 linearly interpolate with respect to time. The scaled plots 
-\ref{fig:na_plot_KOR} and
-\ref{fig:na_plot_USA} do appear to be mostly linear, so we interpolated. 
-In this plots one can observe that $debt_USA$ is decreasing; this is
-because $debt$ carries a sign. The value of $debt_USA$ on
+Figure~\ref{fig:na_plot_KOR} and
+Figure~\ref{fig:na_plot_USA} do appear to be mostly linear, so we interpolated. 
+In this plots one can observe that \texttt{debt\_USA} is decreasing; this is
+because \texttt{debt} carries a sign. The value of \texttt{debt\_USA} on
 
@@
 Figure~\ref{fig:correlation} shows a histogram of all the pairwise sample
-correlations for the original data set- they are
+correlations for the original data set- they are highly correlated. In
+fact, 30 \% of the columns have absolute sample correlation greater than 0.9.
 
@@
 data frame from the original data by taking the difference between each
-subsequent row. More formally, if our original design matrix $X$ is $n
+subsequent row (monthly difference). 
+More formally, if our original design matrix $X$ is $n
 \times p$ then we can define a new $n - 1 \times p$ matrix $\tilde{X}$ by
@@
 
-This approach did indeed eliminate the most obvious patterns, but 
-sacrifices interpretability
+As can be observed in Figure~\ref{fig:diff_scatterplot},
+this approach did indeed eliminate the most obvious patterns, but 
+at the expense of interpretability. We attempted regression on this
+transformed data, but were unsuccessful.
 
-Our first step was to separate the data into training and validation sets,
-and only
+\subsection{Ratios}
 
-Figure~\ref{fig:na_plot_KOR} shows the variables which were interpolated.
+Because exchange rate is a ratio of Korean currency / US dollar we decided
+to make transformed variables consisting of the Korean economic indicator
+divided by the corresponding US indicator. We also included
+\texttt{inflation} untransformed after some preliminary analysis 
+indicated that it was
+important. \texttt{Date} was intentionally left out of the model in order
+to be able to interpret the model in terms of the other parametrs. However,
+the effects of time are still present through the variables that are linear
+over time.
+
+\subsection{Analysis}
+
+Our first step when fitting the model was to separate the data into
+training and validation sets. We checked to make sure that the variables
+looked approximately the same in both training and validation sets.
+
+We used forward stepwise selection to choose a model. 
+Based on the criteria from the output of the \texttt{leaps regsubsets}
+in the appendix we chose to explore a
+model with an intercept and 5 predictor variables. The one with 5
+predictors had the lowest BIC of all the models, and the Mallow's CP at
+5.39 was significantly better than the model with 4 predictors which had a
+CP value of 13.91. Since 5.39 is close to 5 and 13.91 is pretty far from 4
+we conclude that the model with 5 variables is correct, meaning that it has
+little bias.
+
+The variables under further investigation are:
+\texttt{gdp}, \texttt{interest\_rate}, \texttt{inflation\_KOR}, 
+\texttt{inflation\_USA}, and \texttt{govspending}.
+
+The box - cox transformation in figure~\ref{fig:boxcox} suggests a
+transformation of $1 / x^2$, but this extreme and is due to the outliers.
+We used the multiplicative inverse transformation and examined the
+residual plots figure~\ref{fig:cooks} to see why this is the case.
+
+We see
+that the points 113 and 117 are the primary outlying cases. Looking up the
+corresponding dates reveals that these are the points occuring during
+October 2008 and February 2009, during the subprime mortgage crisis
+which shook the global economy. There are no points between 113 and 117
+because the test train split didn't select them. In the validation set we
+discovered similar results for this period. The subprime mortgage crisis is
+the reason that they are outliers.
 
@@
 
+Our use of dynamic data caused trouble because the data was a little too
+dynamic.
+
 \newpage
 \section{Appendix}
+
+Model Selection Output:
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ & sse & bic & r2 & r2a & cp \\
+  \hline
+1 & 0.00 & -46.47 & 0.37 & 0.37 & 411.04 \\
+  2 & 0.00 & -178.62 & 0.80 & 0.79 & 55.28 \\
+  3 & 0.00 & -185.95 & 0.82 & 0.81 & 41.04 \\
+  4 & 0.00 & -206.51 & 0.85 & 0.85 & 13.91 \\
+  5 & 0.00 & -212.36 & 0.86 & 0.86 & 5.39 \\
+  6 & 0.00 & -210.43 & 0.87 & 0.86 & 4.72 \\
+  7 & 0.00 & -207.14 & 0.87 & 0.86 & 5.33 \\
+  8 & 0.00 & -202.47 & 0.87 & 0.86 & 7.22 \\
+  9 & 0.00 & -197.85 & 0.87 & 0.86 & 9.06 \\
+  10 & 0.00 & -193.12 & 0.87 & 0.86 & 11.00 \\
+   \hline
+\end{tabular}
+\end{table}
 
@@
 
+\begin{figure}
+  \centering
+    \includegraphics{diff_scatterplot.pdf}
+  \caption{Scatterplot of data after row difference transformation}
+  \label{fig:diff_scatterplot}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics{boxcox.pdf}
+  \caption{Box cox procedure on training set- affected by outliers}
+  \label{fig:boxcox}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics{cooks.pdf}
+  \caption{Cooks distance for residuals of training data}
+  \label{fig:cooks}
+\end{figure}
+
 \newpage

### 2014-12-13 analysis2.R

@@ -1,0 +1,107 @@
+# Final analysis to produce report
+
+library(MASS)
+library(leaps)
+library(xtable)
+
+source('functions.R')
+load('country.Rda')
+
+
+# Try adding KOR / USD variables
+ratio = with(country, 
+             data.frame(exchange_rate = exchange_rate
+                        , gdp = gdp_KOR / gdp_USA
+                        , unemployment = unemployment_KOR / unemployment_USA
+                        , debt = debt_KOR / debt_USA
+                        , imports = imports_KOR / imports_USA
+                        , exports = exports_KOR / exports_USA
+                        , interest_rate = interest_rate_KOR / interest_rate_USA
+                        , inflation_KOR = inflation_KOR
+                        , inflation_USA = inflation_USA
+                        , cpi = cpi_KOR / cpi_USA
+                        , govspending = govspending_KOR / govspending_USA
+                        ))
+
+# test train split
+set.seed(978)
+a = ttsplit(ratio, 1/3)
+tr = a[[1]]
+test = a[[2]]
+dim(tr)
+head(tr)
+
+# Stepwise regression
+fit6 = regsubsets(I(1 / exchange_rate) ~ ., data=tr, nvmax=11)
+sfit6 = summary(fit6)
+
+modelstats = data.frame(
+    p = 2:11,
+    sse = sfit6$rss,
+    bic = sfit6$bic,
+    r2 = sfit6$rsq,
+    r2a = sfit6$adjr2,
+    cp = sfit6$cp)
+
+modelstats
+xtable(modelstats)
+
+finalvars = sfit6$which[5, ]
+
+tdata = tr[, sfit6$which[5, ]]
+dim(tdata)
+names(tdata)
+
+fit7 = lm(I(1 / exchange_rate) ~ ., data=tdata)
+summary(fit7)
+pdf('boxcox.pdf')
+boxcox(exchange_rate ~ ., data=tdata)
+dev.off()
+
+qqnorm(fit7$residuals)
+# This looks more reasonable
+pdf('cooks.pdf')
+plot(fit7, which=4)
+dev.off()
+# 
+
+infm = influence.measures(fit7)
+summary(infm)
+
+# DFFIT is a little large. Shouldn't be bigger than this:
+2 * sqrt(6 / 121)
+# 113 and 117 are our big ones
+# What happened then?
+country[c(113, 117), 'Date']
+# Prime mortgage rate crash!!
+
[…]

### 2014-12-13 na_plot_KOR.R

@@ -16,25 +16,29 @@ na_plot = function(dframe, filename){
     scaled = data.frame(Date = truncd$Date, scale(truncd[, which(needimpute)]))
     scaled_long = reshape2::melt(scaled, id = 'Date')
 
-    ggplot(data=scaled_long[complete.cases(scaled_long), ]
+    p = ggplot(data=scaled_long[complete.cases(scaled_long), ]
            , aes(x=Date, y=value, color=variable)) + 
            geom_line()
 
-    ggsave(filename)
+    return(p)
 }
 
+p = na_plot(KOR)
+ggsave('na_plot_KOR2.pdf', height = 4, units = 'in')
 
-load('country.Rda')
-
-imputed = c('gdp_KOR', 'gdp_deflator_KOR', 'debt_KOR', 
-            'interest_rate_KOR', 'govspending_KOR')
-
-scaled = data.frame(Date = KOR$Date, scale(KOR[, imputed]))
-
-scaled_long = reshape2::melt(scaled, id = 'Date')
-
-ggplot(data=scaled_long[complete.cases(scaled_long), ]
-       , aes(x=Date, y=value, color=variable)) + 
-       geom_line()
-
-ggsave('na_plot_KOR.pdf', height = 4, units = 'in')
+#
+#
+#load('country.Rda')
+#
+#imputed = c('gdp_KOR', 'gdp_deflator_KOR', 'debt_KOR', 
+#            'interest_rate_KOR', 'govspending_KOR')
+#
+#scaled = data.frame(Date = KOR$Date, scale(KOR[, imputed]))
+#
+#scaled_long = reshape2::melt(scaled, id = 'Date')
+#
+#ggplot(data=scaled_long[complete.cases(scaled_long), ]
+#       , aes(x=Date, y=value, color=variable)) + 
+#       geom_line()
+#
+#ggsave('na_plot_KOR.pdf', height = 4, units = 'in')

### 2014-12-14 paper.tex
centered model

@@
 2014. In 1997 the IMF crisis in Korea perturbed the economic indicators,
-but in looking at the data we can see that it has been stable since 1999.
+but they have been mostly stable since 1999.
 
@@
         government. Units: local currency
-    \item {\tt political party} Categorical variable for the political
-party of the president.
 \end{itemize}
@@
 
-\subsection{Reproducibility}
+Through a little engineering the data analysis process
+can be made transparent, automated, extensible, and fully reproducible.
+For more information on the techniques please refer to the section on
+Reproducibility in the appendix.
+
+\subsection{Exploratory Data Analysis}
+
+We started out with some general summary statistics and plotting. The first
+issue was missing data, which we had anticipated. While most of the
+variables were reported monthly, some were reported on a quarterly or
+annual basis. (Figure~\ref{fig:na_plot_KOR})
+Our simple approach for dealing with this was to
+make a plot of these variables over time.
+If they are approximately linear then we could 
+linearly interpolate with respect to time. The scaled plots 
+Figure~\ref{fig:na_plot_KOR} and
+Figure~\ref{fig:na_plot_USA} do appear to be mostly linear, so we interpolated. 
+In this plots one can observe that \texttt{debt\_USA} is decreasing; this is
+because \texttt{debt} carries a sign.
+
+When beginning this project we knew that the time dependency and
+multicollinearity would be the
+biggest obstacles, since in this case the assumption that the errors
+$\epsilon_i \sim N(0, \sigma^2)$ does not hold.
+Figure~\ref{fig:scatterplot} shows what the original data looks like.
+We observe clear patterns emerging, with some even approaching a
+continuous mathematical functional relationship- ie the patterns
+between \texttt{gdp}, \texttt{gdp\_deflator}, and \texttt{debt} for Korea.
+Figure~\ref{fig:correlation} shows a histogram of all the pairwise sample
+correlations for the original data set- they are highly correlated. In
+fact, 30 \% of the columns have absolute sample correlation greater than 0.9.
+
+Our first approach for overcoming this time dependence was to create a new
+data frame from the original data by taking the difference between each
+subsequent row (monthly difference). 
+More formally, if our original design matrix $X$ is $n
+\times p$ then we can define a new $n - 1 \times p$ matrix $\tilde{X}$ by
+setting
+\[
+    \tilde{X_{i,*}} = X_{(i + 1), *} - X_{i, *}
+\]
+for $i = 1, 2, \dots, n-1$ where $X_{i, *}$ represents the $i$th row of
+$X$.
+
+As can be observed in Figure~\ref{fig:diff_scatterplot},
+this approach did indeed eliminate the most obvious patterns, but 
+at the expense of interpretability. We attempted regression on this
+transformed data, but were unsuccessful.
+
+\subsection{Ratios}
+
+Because exchange rate is a ratio of Korean currency / US dollar we decided
+to make transformed variables consisting of the Korean economic indicator
+divided by the corresponding US indicator. We also included
+\texttt{inflation} untransformed since the US has value 0 for some time
+periods, meaning that this transformation is not defined.
+\texttt{Date} was intentionally left out of the model in order
+to be able to interpret the model in terms of the other parametrs. However,
+the effects of time are still present through the variables that are linear
+over time.
+
+\subsection{Analysis}
+
+Our first step when fitting the model was to separate the data into
+training and validation sets. We checked to make sure that the variables
+looked approximately the same in both training and validation sets.
+
+We used forward stepwise selection to choose a model. 
+Based on the criteria from the output of the \texttt{leaps regsubsets}
+in the appendix we chose to explore a
+model with an intercept and 5 predictor variables. The one with 5
+predictors had the lowest BIC of all the models, and the Mallow's CP at
+5.39 was significantly better than the model with 4 predictors which had a
+CP value of 13.91. Since 5.39 is close to 5 and 13.91 is pretty far from 4
+we conclude that the model with 5 variables is correct, meaning that it has
+little bias.
+
+The variables under further investigation are:
+\texttt{gdp}, \texttt{interest\_rate}, \texttt{inflation\_KOR}, 
+\texttt{inflation\_USA}, and \texttt{govspending}. (Model 1)
+
+The box - cox transformation in figure~\ref{fig:boxcox} suggests a
+transformation of $1 / x^2$, but this is extreme and is due to the
+outliers. Instead we used the multiplicative inverse transformation on the
+\texttt{exchange\_rate}, which has the appealing property of maintaining
+interpretability; instead of Korean Won / USD we have USD / Korean Won.
+
+The outliers can be readily detected in the diagnostic plots 
+figure~\ref{fig:cooks}. We see
+that the points 113 and 117 are the primary outlying cases, which have
+absolute DFFIT values of 1.16 and 0.89. This is much bigger than the cutoff
+of $2 / \sqrt{p / n} = 0.445$. Cook's distance is large as well.
+Looking up the
+corresponding dates reveals that these are the points occuring during
+October 2008 and February 2009, during the subprime mortgage crisis
+which shook the global economy. There are no points between 113 and 117
+because the test train split didn't select them. In the validation set we
+discovered similar results for this period(114, 118). The subprime mortgage
+crisis is the reason that they are outliers. So, we took off the outliers
+from our data set to get model 1 fit~\ref{m1:train}. We also had trouble
+with recent dates having large leverage values of around 0.16, which is
+greater than the threshold of $2 * p / n = 0.08$.
+
+When checking this final model with the validation set, then the
+\texttt{govspending} variable is not
+significant (P-value:0.312)~\ref{m1:test}. The added-variable plot in
+figure~\ref{fig:added} indicates \texttt{govspending} contains no
+additional information useful for predicting (1/exchage\_rate) beyond the
+other variable, so it is not helpful to add \texttt{govspending} to our
+model. Thus, we decide the our final variables are \texttt{gdp},
+\texttt{interest\_rate}, \texttt{inflation\_KOR}, and
+\texttt{inflation\_USA}~\ref{finalmodel}. The corresponding equation is:
+
+$\frac{1}{exchange\_rate} = 3.868323e^{-4} + 9.692157e^{-3} * gdp
+-1.289123e^{-5} *interest\_rate -2.573026e^{-5} * inflation\_KOR +
+2.135945e^{-5}  * inflation\_USA$
+
+For this model we have $R^2 = 0.85, MSE = 1.28^{-9}$.
+
+\subsection{Centering and Signs}
+
+The model above is difficult to interpret, so we decided to center the data
+to minimize multicollinearity and get more accessible numbers. After a
+similar analysis process as described above we found that the final model
+for the centered data has the same variables as well as \texttt{export}.
+
+common variables and export variable. Interesting part of this trial is
+about outliers. Its outliers are showing up 2000 finanical crisis rather
+than the subprime mortage crisis (it fitted well)~\ref{centered}.
+
+We used Bonferroni's method to construct simultaneous 95\% confidence intervals
+for each of the 5 variables selected in the centered model.
+
+\section{Conclusions and Discussion}
+
+Our use of dynamic data caused trouble because the data was a little too
+dynamic.
+
+Our model: $\frac{USA dollar Worthy}{KOREA Won Worthy} \sim \frac{KOREA gdp}{USA GDP} - \frac{KOREA Interest Rate}{USA Interest Rate} -  Korea inflation + USA inflation$, means when relative Korea gdp over US gdp is increasing, the won per a dollar is decreasing.
+The Korea Interest Rate is relatively increased over the USA Interest rate(higher interest rate means higher liquidity of money in markets), which bring the won per a dollar is increasing.
+Also, the signs of Korea inflation and USA inflation are different, which is very reliable because we are predicting the $\frac{USA dollar Worthy}{KOREA Won Worthy}$.
+
+Commonly, we can think if you want to exchage your currency to other foriegn currency, it would better off to do it when your country's relative gdp is greater than the foriegn, when your country's relative interest rate is less , when your contry is not on inflation, or when the foriegn country is on inflation situation.
+We have noticed exchange rates are dependent on previous exchage rates(time series) if we are available to do mutltivariate time series.
+
+\newpage
+
+\section{Appendix - Reproducibility}
 
@@
 
-Modern high level programming languages like R provide incredible tools. By
-using their capabilities we can
-
-\subsection{Exploratory Data Analysis}
-
-We started out with some general summary statistics and plotting. The first
-issue was missing data, which we had anticipated. While most of the
-variables were reported monthly, \texttt{debt} was reported on a quarterly
-basis and 
-Our simple approach for dealing with this was to
-make a plot of these variables over time.
-If they are approximately linear then we could 
-linearly interpolate with respect to time. The scaled plots 
-Figure~\ref{fig:na_plot_KOR} and
-Figure~\ref{fig:na_plot_USA} do appear to be mostly linear, so we interpolated. 
-In this plots one can observe that \texttt{debt\_USA} is decreasing; this is
-because \texttt{debt} carries a sign. The value of \texttt{debt\_USA} on
-
-When beginning this project we knew that the time dependency and
-multicollinearity would be the
-biggest obstacles, since in this case the assumption that the errors
-$\epsilon_i \sim N(0, \sigma^2)$ does not hold.
-Figure~\ref{fig:scatterplot} shows what the original data looks like.
-We observe clear patterns emerging, with some even approaching a
-continuous mathematical functional relationship- ie the patterns
-between \texttt{gdp}, \texttt{gdp\_deflator}, and \texttt{debt} for Korea.
-Figure~\ref{fig:correlation} shows a histogram of all the pairwise sample
-correlations for the original data set- they are highly correlated. In
-fact, 30 \% of the columns have absolute sample correlation greater than 0.9.
-
-Our first approach for overcoming this time dependence was to create a new
-data frame from the original data by taking the difference between each
-subsequent row (monthly difference). 
-More formally, if our original design matrix $X$ is $n
-\times p$ then we can define a new $n - 1 \times p$ matrix $\tilde{X}$ by
-setting
-\[
-    \tilde{X_{i,*}} = X_{(i + 1), *} - X_{i, *}
-\]
-for $i = 1, 2, \dots, n-1$ where $X_{i, *}$ represents the $i$th row of
-$X$.
-
-As can be observed in Figure~\ref{fig:diff_scatterplot},
-this approach did indeed eliminate the most obvious patterns, but 
-at the expense of interpretability. We attempted regression on this
-transformed data, but were unsuccessful.
-
-\subsection{Ratios}
-
-Because exchange rate is a ratio of Korean currency / US dollar we decided
-to make transformed variables consisting of the Korean economic indicator
-divided by the corresponding US indicator. We also included
-\texttt{inflation} untransformed after some preliminary analysis 
-indicated that it was
-important. \texttt{Date} was intentionally left out of the model in order
-to be able to interpret the model in terms of the other parametrs. However,
-the effects of time are still present through the variables that are linear
-over time.
-
-\subsection{Analysis}
-
-Our first step when fitting the model was to separate the data into
-training and validation sets. We checked to make sure that the variables
-looked approximately the same in both training and validation sets.
-
-We used forward stepwise selection to choose a model. 
-Based on the criteria from the output of the \texttt{leaps regsubsets}
-in the appendix we chose to explore a
-model with an intercept and 5 predictor variables. The one with 5
-predictors had the lowest BIC of all the models, and the Mallow's CP at
-5.39 was significantly better than the model with 4 predictors which had a
-CP value of 13.91. Since 5.39 is close to 5 and 13.91 is pretty far from 4
-we conclude that the model with 5 variables is correct, meaning that it has
-little bias.
-
-The variables under further investigation are:
-\texttt{gdp}, \texttt{interest\_rate}, \texttt{inflation\_KOR}, 
-\texttt{inflation\_USA}, and \texttt{govspending}.
-
-The box - cox transformation in figure~\ref{fig:boxcox} suggests a
-transformation of $1 / x^2$, but this extreme and is due to the outliers.
-We used the multiplicative inverse transformation and examined the
-residual plots figure~\ref{fig:cooks} to see why this is the case.
-
-We see
-that the points 113 and 117 are the primary outlying cases. Looking up the
-corresponding dates reveals that these are the points occuring during
-October 2008 and February 2009, during the subprime mortgage crisis
-which shook the global economy. There are no points between 113 and 117
-because the test train split didn't select them. In the validation set we
-discovered similar results for this period. The subprime mortgage crisis is
-the reason that they are outliers.
-
-Attempting to reference \cite{lamport94}.
-
-\section{Conclusions and Discussion}
-
-Our use of dynamic data caused trouble because the data was a little too
-dynamic.
-
-\newpage
-\section{Appendix}
-
-Model Selection Output:
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrr}
-  \hline
- & sse & bic & r2 & r2a & cp \\
-  \hline
-1 & 0.00 & -46.47 & 0.37 & 0.37 & 411.04 \\
-  2 & 0.00 & -178.62 & 0.80 & 0.79 & 55.28 \\
-  3 & 0.00 & -185.95 & 0.82 & 0.81 & 41.04 \\
-  4 & 0.00 & -206.51 & 0.85 & 0.85 & 13.91 \\
-  5 & 0.00 & -212.36 & 0.86 & 0.86 & 5.39 \\
-  6 & 0.00 & -210.43 & 0.87 & 0.86 & 4.72 \\
-  7 & 0.00 & -207.14 & 0.87 & 0.86 & 5.33 \\
-  8 & 0.00 & -202.47 & 0.87 & 0.86 & 7.22 \\
-  9 & 0.00 & -197.85 & 0.87 & 0.86 & 9.06 \\
-  10 & 0.00 & -193.12 & 0.87 & 0.86 & 11.00 \\
-   \hline
-\end{tabular}
-\end{table}
+\section{Appendix 1}
 
@@
     \includegraphics{na_plot_USA.pdf}
-  \caption{Scaled variables with NA values}
+  \caption{This figure illustrates the perils of depending on live data. US debt was
+formerly measured as a negative number, which is why it was decreasing. But
+midway through our analysis the source began to return a huge positive
+number for the most recent dates.}
+
   \label{fig:na_plot_USA}
 \end{figure}
-
 \begin{figure}
@@
 
+\begin{figure}
+	\centering
+	\includegraphics{Addedvariable.pdf}
+	\caption{Added-Variable Plot for Government Spending}
+	\label{fig:added}
+\end{figure}
+
 \newpage
 
-\begin{thebibliography}{9}
+\section{Appendix 2}
 
-\bibitem{lamport94}
-  Leslie Lamport,
-  \emph{\LaTeX: a document preparation system}.
-  Addison Wesley, Massachusetts,
-  2nd edition,
-  1994.
+\subsection{Model Selection Output}
+\begin{table}[ht]
+	\centering
+	\begin{tabular}{rrrrrr}
+		\hline
+		& sse & bic & r2 & r2a & cp \\
+		\hline
+		1 & 0.00 & -46.47 & 0.37 & 0.37 & 411.04 \\
+		2 & 0.00 & -178.62 & 0.80 & 0.79 & 55.28 \\
+		3 & 0.00 & -185.95 & 0.82 & 0.81 & 41.04 \\
+		4 & 0.00 & -206.51 & 0.85 & 0.85 & 13.91 \\
+		5 & 0.00 & -212.36 & 0.86 & 0.86 & 5.39 \\
+		6 & 0.00 & -210.43 & 0.87 & 0.86 & 4.72 \\
+		7 & 0.00 & -207.14 & 0.87 & 0.86 & 5.33 \\
+		8 & 0.00 & -202.47 & 0.87 & 0.86 & 7.22 \\
+		9 & 0.00 & -197.85 & 0.87 & 0.86 & 9.06 \\
+		10 & 0.00 & -193.12 & 0.87 & 0.86 & 11.00 \\
+		\hline
+	\end{tabular}
+\end{table}
 
-\end{thebibliography}
+\subsection{Influence Measures}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrrrrrr}
+  \hline
+ & dfb.1\_ & dfb.gdp & dfb.int\_ & dfb.i\_KO & dfb.i\_US & dfb.gvsp & dffit & cov.r & cook.d & hat \\
+  \hline
+110 & -0.15 & -0.04 & -0.09 & 0.18 & 0.06 & 0.11 & 0.25 & 1.16 & 0.01 & 0.12 \\
+  113 & 0.67 & 0.18 & 0.43 & -0.93 & 0.00 & -0.51 & -1.17 & 0.30 & 0.18 & 0.05 \\
+  117 & 0.24 & 0.01 & 0.21 & -0.65 & 0.70 & -0.18 & -0.89 & 0.67 & 0.12 & 0.07 \\
+  122 & -0.02 & -0.02 & 0.01 & -0.00 & 0.07 & 0.01 & -0.08 & 1.18 & 0.00 & 0.11 \\
+  173 & -0.03 & -0.03 & -0.03 & -0.00 & -0.01 & 0.05 & 0.08 & 1.16 & 0.00 & 0.09 \\
+  176 & -0.05 & -0.05 & -0.05 & 0.01 & 0.00 & 0.09 & 0.11 & 1.24 & 0.00 & 0.16 \\
+  177 & -0.06 & -0.05 & -0.05 & 0.01 & -0.01 & 0.09 & 0.11 & 1.24 & 0.00 & 0.15 \\
+  178 & -0.06 & -0.05 & -0.05 & 0.02 & -0.00 & 0.10 & 0.11 & 1.24 & 0.00 & 0.16 \\
+  179 & -0.21 & -0.20 & -0.18 & 0.06 & 0.02 & 0.33 & 0.38 & 1.20 & 0.02 & 0.16 \\
+  180 & -0.37 & -0.33 & -0.30 & 0.13 & 0.04 & 0.57 & 0.64 & 1.12 & 0.07 & 0.16 \\
+   \hline
+\end{tabular}
+\end{table}
+
+\subsubsection{Model1 summary output (Train Data)\label{m1:train}}
+\begin{verbatim}
+Call:
+lm(formula = I(1/exchange_rate) ~ ., data = tdata8)
+
+Residuals:
+Min         1Q     Median         3Q        Max 
+-7.807e-05 -1.661e-05 -3.530e-07  2.128e-05  6.973e-05
+
+Coefficients:
+               Estimate   Std. Error t value Pr(>|t|)    
+(Intercept)    3.901e-04  2.851e-05  13.683  < 2e-16 ***
+gdp            1.130e-02  5.522e-04  20.471  < 2e-16 ***
+interest_rate -1.211e-05  1.085e-06 -11.158  < 2e-16 ***
+inflation_KOR -1.877e-05  3.334e-06  -5.630 1.35e-07 ***
+inflation_USA  1.529e-05  2.574e-06   5.943 3.23e-08 ***
+govspending   -9.170e-06  3.423e-06  -2.679   0.0085 ** 
+---
+Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+
+Residual standard error: 2.863e-05 on 112 degrees of freedom
+Multiple R-squared:  0.9034,	Adjusted R-squared:  0.8991 
+F-statistic: 209.4 on 5 and 112 DF,  p-value: < 2.2e-16
+\end{verbatim}
+
+\subsubsection{Model1 summary output (Validation Data)\label{m1:test}}
+
+\begin{verbatim}
+Call:
+lm(formula = I(1/exchange_rate) ~ ., data = test99)
+
+Residuals:
+Min         1Q     Median         3Q        Max 
+-6.962e-05 -1.878e-05 -2.636e-06  1.988e-05  4.937e-05
+
+Coefficients:
+               Estimate   Std. Error t value Pr(>|t|)    
+(Intercept)    4.038e-04  3.592e-05  11.244 1.56e-15 ***
+gdp            1.001e-02  7.838e-04  12.769  < 2e-16 ***
+interest_rate -1.163e-05  1.601e-06  -7.268 1.85e-09 ***
+inflation_KOR -2.808e-05  3.749e-06  -7.492 8.15e-10 ***
+inflation_USA  2.315e-05  4.847e-06   4.776 1.50e-05 ***
+govspending   -3.640e-06  4.246e-06  -0.857    0.395    
+---
+Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+
+Residual standard error: 2.742e-05 on 52 degrees of freedom
+Multiple R-squared:  0.9161,	Adjusted R-squared:  0.908 
+F-statistic: 113.5 on 5 and 52 DF,  p-value: < 2.2e-16
+\end{verbatim}
+
+\subsubsection{Summary table of Final model}
+
+\label{finalmodel}
+\begin{verbatim}
+Call:
+lm(formula = I(1/exchange_rate) ~ gdp + interest_rate + inflation_KOR + 
+inflation_USA, data = alldata)
+
+Residuals:
+Min         1Q     Median         3Q        Max 
+-1.631e-04 -1.924e-05  4.079e-06  2.177e-05  8.537e-05
+
+Coefficients:
+               Estimate   Std. Error t value Pr(>|t|)    
+(Intercept)    3.868e-04  2.285e-05  16.929  < 2e-16 ***
+gdp            9.692e-03  4.007e-04  24.190  < 2e-16 ***
+interest_rate -1.289e-05  8.636e-07 -14.927  < 2e-16 ***
+inflation_KOR -2.573e-05  2.637e-06  -9.758  < 2e-16 ***
+inflation_USA  2.136e-05  2.646e-06   8.072 1.04e-13 ***
+---
+Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+
+Residual standard error: 3.58e-05 on 176 degrees of freedom
+Multiple R-squared:  0.8555,	Adjusted R-squared:  0.8522 
+F-statistic: 260.5 on 4 and 176 DF,  p-value: < 2.2e-16
+\end{verbatim}
+
+\subsubsection{Final model Summary Output(Centered data)}
+\label{centered}
+
+\begin{verbatim}
+Call:
+lm(formula = exchange_rate ~ gdp + exports + interest_rate + 
+inflation_KOR + inflation_USA, data = sdata_n)
+
+Residuals:
+Min       1Q   Median       3Q      Max 
+-0.72512 -0.22635  0.00475  0.16783  0.94453
+
+Coefficients:
+              Estimate    Std. Error t value Pr(>|t|)    
+(Intercept)   -0.03923    0.02417  -1.623    0.106    
+gdp           -0.68565    0.04872 -14.073  < 2e-16 ***
+exports       -0.24752    0.05083  -4.869 2.54e-06 ***
+interest_rate  0.59072    0.03611  16.361  < 2e-16 ***
+inflation_KOR  0.28550    0.02762  10.337  < 2e-16 ***
+inflation_USA -0.29924    0.03197  -9.361  < 2e-16 ***
+---
+Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+
+Residual standard error: 0.3211 on 171 degrees of freedom
+Multiple R-squared:  0.8868,	Adjusted R-squared:  0.8835 
+F-statistic: 267.9 on 5 and 171 DF,  p-value: < 2.2e-16
+\end{verbatim}
+
+\subsubsection{Bonferroni's confidence intervals on centered coefficients}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrr}
+  \hline
+ & lower & upper \\
+  \hline
+(Intercept) & -0.10 & 0.02 \\
+  gdp & -0.80 & -0.57 \\
+  exports & -0.37 & -0.13 \\
+  interest\_rate & 0.51 & 0.68 \\
+  inflation\_KOR & 0.22 & 0.35 \\
+  inflation\_USA & -0.37 & -0.22 \\
+   \hline
+\end{tabular}
+\end{table}
 

### 2014-12-14 analysis2.R

@@ -66,10 +66,12 @@ dev.off()
 # 
 
 infm = influence.measures(fit7)
-summary(infm)
+xtable(summary(infm))
 
-# DFFIT is a little large. Shouldn't be bigger than this:
+# DFFIT is a little large at -1.16 and -0.89. Shouldn't be bigger than this:
 2 * sqrt(6 / 121)
+# cooks distance
+
 # 113 and 117 are our big ones
 # What happened then?
 country[c(113, 117), 'Date']
@@ -98,10 +100,24 @@ plot.lm(fit9, which=4)
 coef(fit8)
 coef(fit9)
 
-form10 = exchange_rate ~ gdp + interest_rate + inflation_KOR + inflation_USA
+form10 = I(1 / exchange_rate) ~ gdp + interest_rate + inflation_KOR + inflation_USA
 fit10 = lm(form10, data= tdata8)
 # Do we still need to do inverse transform if outliers are gone? 
 boxcox(fit10)
 # Still calls for the inverse transform even with the outliers removed.
 summary(fit10)
 
+fit11 = lm(form10, data= ratio)
+summary(fit11)
+xtable(confint(fit11), digits=-3)
+
+nooutlier = ratio[-(110:120), ]
+# With outliers gone
+fit12 = lm(form10, data= nooutlier)
+summary(fit12)
+xtable(confint(fit12))
+
+# Centering
+data13 = data.frame(scale(ratio))
+fit13 = lm(exchange_rate ~ gdp + interest_rate + inflation_KOR + inflation_USA, data=data13)
+summary(fit13)

### 2014-12-14 exploratory+.R

@@ -1,0 +1,376 @@
+library(ggplot2)
+
+# Exploratory data analysis
+load('country.Rda')
+source('functions.R')
+
+# Why is debt going down for the US? Because it's a huge negative number!
+# Plot it and look at the range in order to see this.
+
+# Exchange rate over time
+qplot(Date, exchange_rate, data = country, ylab = 'Won / Dollar')
+ggsave('exchange_rate.pdf', width = 6, height = 8, units = 'in')
+
+# What's the max difference?
+r = range(country$exchange_rate)
+r[2] / r[1]
+
+# Will it fit?
+fit1 = lm(exchange_rate ~ ., data = trainset[, -1])
+summary(fit1)
+plot(fit1)
+
+# Intentionally Remove the effects of time
+timefit = lm(exchange_rate ~ Date, data=trainset)
+plot(timefit)
+# This is a terrible fit - exchange rate is not linear over time.
+# Anyways, it would be difficult to fully remove time
+
+country_notime = cbind(exchange_notime = trainset$exchange_rate - predict(timefit),
+                       trainset[, -c(1, 2)])
+
+fit2 = lm(exchange_notime ~ ., data = country_notime)
+summary(fit2)
+plot(fit2)
+
+
+# Thu Dec 11 14:18:14 PST 2014
+
+set.seed(123)
+
+# Drop the date column
+nodate = country[, names(country) != 'Date']
+dim(nodate)
+
+# Exploratory data analysis
+pdf('scatterplot.pdf')
+plot(nodate[, 1:5])
+dev.off()
+
+plot(nodate[, 1:5])
+# Time dependent patterns are very obvious here
+# We want to be seeing a cloud
+# Who put these snakes in my data?!!
+plot(nodate[, c(1, 6:10)])
+plot(nodate[, c(1, 11:16)])
+plot(nodate[, c(1, 17:21)])
+
+# Suppose we pick out the variables where the patterns between
+# exchange_rate and the other variables are least obvious?
+# We can see them from the plots above.
+nopattern = c('exchange_rate', 'unemployment_KOR', 'interest_rate_KOR',
+              'inflation_KOR', 'interest_rate_USA', 'inflation_USA')
+
+# What happens if we just regress on these?
+fitnopattern = lm(exchange_rate ~ ., data = nodate[, nopattern])
[…]

### 2014-12-14 paper.tex
abstract

@@
 \begin{abstract}
+
     We analyze the exchange rate between the United States and South Korea
-    (hereafter Korea) using monthly country level economic data.
+    (hereafter Korea) using monthly country level economic data and find
+    that ratios of economic indicators are useful in confirming economic
+    reasoning about the relative strengths of the two national economies.
+
 \end{abstract}
@@
 absolute DFFIT values of 1.16 and 0.89. This is much bigger than the cutoff
-of $2 / \sqrt{p / n} = 0.445$. Cook's distance is large as well.
+of $2 \sqrt{p / n} = 0.445$. Cook's distance is large as well.
 Looking up the
@@
 The model above is difficult to interpret, so we decided to center the data
-to minimize multicollinearity and get more accessible numbers. After a
-similar analysis process as described above we found that the final model
-for the centered data has the same variables as well as \texttt{export}.
-
-common variables and export variable. Interesting part of this trial is
-about outliers. Its outliers are showing up 2000 finanical crisis rather
-than the subprime mortage crisis (it fitted well)~\ref{centered}.
+to minimize multicollinearity and instead focus on the sign of the
+coefficients. After a
+similar analysis process we found that the final model
+for the centered data has the same variables as well as \texttt{exports}.
+The interesting part here were again the outliers, which this time also included
+a point from the 2000 financial crisis (see \ref{centered} in the appendix). This fit came out
+quite well, and the corresponding diagnostic plots can be seen in 
+Figure \ref{fig:final_residual}.
 
 We used Bonferroni's method to construct simultaneous 95\% confidence intervals
-for each of the 5 variables selected in the centered model.
+for each of the 5 variables selected in the centered model, producing the
+following intervals:
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrr}
+  \hline
+ & lower & upper \\
+  \hline
+  gdp & -0.80 & -0.57 \\
+  exports & -0.37 & -0.13 \\
+  interest\_rate & 0.51 & 0.68 \\
+  inflation\_KOR & 0.22 & 0.35 \\
+  inflation\_USA & -0.37 & -0.22 \\
+   \hline
+\end{tabular}
+\end{table}
 
@@
 
-Our use of dynamic data caused trouble because the data was a little too
-dynamic.
+Let $Y$ be the exchange rate of Korean Won per US Dollar. Let $G, E, R, I$
+represent the value of GDP, exports, interest rate, and inflation, respectively.
+Let $X_K$ be the values for Korea and $X_{US}$ be the ones for the US. Let
+$c_i$ be a nonnegative scalar for scaling.
+The final model becomes
 
-Our model: $\frac{USA dollar Worthy}{KOREA Won Worthy} \sim \frac{KOREA gdp}{USA GDP} - \frac{KOREA Interest Rate}{USA Interest Rate} -  Korea inflation + USA inflation$, means when relative Korea gdp over US gdp is increasing, the won per a dollar is decreasing.
-The Korea Interest Rate is relatively increased over the USA Interest rate(higher interest rate means higher liquidity of money in markets), which bring the won per a dollar is increasing.
-Also, the signs of Korea inflation and USA inflation are different, which is very reliable because we are predicting the $\frac{USA dollar Worthy}{KOREA Won Worthy}$.
+\[
+    Y = -0.69 c_1 \frac{G_{K}}{G_{US}}
+    -0.25 c_2 \frac{E_{K}}{E_{US}}
+    +0.59 c_3 \frac{R_{K}}{R_{US}}
+    +0.29 c_4I_{K} - 0.30 c_5 I_{US} + \epsilon.
+\]
 
-Commonly, we can think if you want to exchage your currency to other foriegn currency, it would better off to do it when your country's relative gdp is greater than the foriegn, when your country's relative interest rate is less , when your contry is not on inflation, or when the foriegn country is on inflation situation.
-We have noticed exchange rates are dependent on previous exchage rates(time series) if we are available to do mutltivariate time series.
+The 95\% confidence interval does not include 0 for any of these variables,
+so we are 95\% confident that these signs are correct. Since the predictor variables
+are scaled or ratios they are unitless. Now the model has a
+natural interpretation. Suppose that the Korean economy grows while the US
+economy remains stagnant. Then the exchange rate should decrease, because
+the Korean Won will have more value relative to the US Dollar.
+From our model we can infer the following:
+
+\begin{enumerate}
+    \item The negative coefficient of $\frac{G_{K}}{G_{US}}$ means that as
+        the GDP of Korea increases relative to that of the US, the exchange
+        rate decreases. 
+    \item The negative coefficient of $\frac{E_{K}}{E_{US}}$ means that if
+        Korea increases its exports relative to the US, the exchange rate
+        decreases.
+    \item The positive coefficient of $I_K$ means that the exchange rate
+        increases as the inflation in Korea gets larger. This is a direct
+        relation that we would expect.
+    \item The negative coefficient of $I_{US}$ means that the exchange rate
+        decreases as the inflation in the US gets larger.
+\end{enumerate}
+
+Common sense and economic reasoning tells us that all of these should be
+true. The data and analysis confirm our intuition. Perhaps with the use of
+time series techniques this analysis could be extended.
+
+Returning to out original questions, the results of this analysis suggest
+that if you want to exchange your currency
+to foreign currency, it would be better to do it when your home country's 
+GDP is good, or the exports are good, or the inflation is relatively lower.
 
@@
 	\label{fig:added}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics{final_residual.pdf}
+  \caption{final residual plots for centered model}
+  \label{fig:final_residual}
 \end{figure}
@@
 
-\subsubsection{Bonferroni's confidence intervals on centered coefficients}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrr}
-  \hline
- & lower & upper \\
-  \hline
-(Intercept) & -0.10 & 0.02 \\
-  gdp & -0.80 & -0.57 \\
-  exports & -0.37 & -0.13 \\
-  interest\_rate & 0.51 & 0.68 \\
-  inflation\_KOR & 0.22 & 0.35 \\
-  inflation\_USA & -0.37 & -0.22 \\
-   \hline
-\end{tabular}
-\end{table}
-
 \end{document}

### 2014-12-14 exploratory+.R

@@ -359,6 +359,10 @@ plot(sfit5$residual, adds$residual) # show definitely negative trend, so shouldn
 
 sfinal = lm(exchange_rate ~ gdp + exports + interest_rate + inflation_KOR + inflation_USA, data = sdata_n)
 summary(sfinal) # all variables are significants
+pdf('final_residual.pdf')
+par(mfrow=c(2,1))
+plot.lm(sfinal, which=c(1,2))
+dev.off()
 
 #Actually this result from scaling data is very interesting. The resason I tried it because MSE is too small number(third decimal point)(all are in small range) in the not scaled data, so I wanted to get more reasonable number, so scale the data and try it again.
 #By scaling effects, the coefficents and MSE become closer to whole numbers, but comparing the two types of outliers(from scaled data and from non-scaled data) are not common. Non-scaled data outlier showes 2007-2008 financial crisis, and scaled data outlier shows 2000 financial crisis. Why they showed up only one financial crisis not both of them??
@@ -367,10 +371,16 @@ summary(sfinal) # all variables are significants
 dim(sdata)
 colMeans(sdata)
 
-sumfinal = summary(sfinal) # all variables are significants
+sumfinal = summary(sfinal) 
+
+coef(sfinal)
+
+sinf = influence.measures(sfinal)
+summary(sinf)
 
 # Bonferroni's for simultaneous 95 percent confidence of 5 coefficients
 multiplier = qt(1 - 0.05 / 5, nrow(sdata) - 5) * sumfinal$coefficients[, 'Std. Error']
 sfinalconf = data.frame(lower = coef(sfinal) - multiplier,
                         upper = coef(sfinal) + multiplier)
+sfinalconf
 xtable(sfinalconf)

### 2014-12-14 paper.tex
merging exploratory files

@@
 
-    \item When is the best time to exchange currency from won (Korean
-        currency) to dollar or vice versa?
+    \item When is the best time to exchange currency from Korean Won
+        to US Dollar or vice versa?
 
@@
 that we've been active on this project every day since
-December 1st. Here's an example of a log entry:
+December 1st, pushed more than 30 commits, and wrote nearly 3000 lines of
+code (including \LaTeX). Here's an example of a log entry:
 

### 2014-12-14 exploratory+.R

@@ -1,386 +1,0 @@
-library(ggplot2)
-
-# Exploratory data analysis
-load('country.Rda')
-source('functions.R')
-
-# Why is debt going down for the US? Because it's a huge negative number!
-# Plot it and look at the range in order to see this.
-
-# Exchange rate over time
-qplot(Date, exchange_rate, data = country, ylab = 'Won / Dollar')
-ggsave('exchange_rate.pdf', width = 6, height = 8, units = 'in')
-
-# What's the max difference?
-r = range(country$exchange_rate)
-r[2] / r[1]
-
-# Will it fit?
-fit1 = lm(exchange_rate ~ ., data = trainset[, -1])
-summary(fit1)
-plot(fit1)
-
-# Intentionally Remove the effects of time
-timefit = lm(exchange_rate ~ Date, data=trainset)
-plot(timefit)
-# This is a terrible fit - exchange rate is not linear over time.
-# Anyways, it would be difficult to fully remove time
-
-country_notime = cbind(exchange_notime = trainset$exchange_rate - predict(timefit),
-                       trainset[, -c(1, 2)])
-
-fit2 = lm(exchange_notime ~ ., data = country_notime)
-summary(fit2)
-plot(fit2)
-
-
-# Thu Dec 11 14:18:14 PST 2014
-
-set.seed(123)
-
-# Drop the date column
-nodate = country[, names(country) != 'Date']
-dim(nodate)
-
-# Exploratory data analysis
-pdf('scatterplot.pdf')
-plot(nodate[, 1:5])
-dev.off()
-
-plot(nodate[, 1:5])
-# Time dependent patterns are very obvious here
-# We want to be seeing a cloud
-# Who put these snakes in my data?!!
-plot(nodate[, c(1, 6:10)])
-plot(nodate[, c(1, 11:16)])
-plot(nodate[, c(1, 17:21)])
-
-# Suppose we pick out the variables where the patterns between
-# exchange_rate and the other variables are least obvious?
-# We can see them from the plots above.
-nopattern = c('exchange_rate', 'unemployment_KOR', 'interest_rate_KOR',
-              'inflation_KOR', 'interest_rate_USA', 'inflation_USA')
-
-# What happens if we just regress on these?
-fitnopattern = lm(exchange_rate ~ ., data = nodate[, nopattern])
[…]

### 2014-12-14 exploratory.R

@@ -234,6 +234,9 @@ summary(infm)
 
 # DFFIT is a little large. Shouldn't be bigger than this:
 2 * sqrt(6 / 121)
+
+df(0.2, 6, 121 - 6)
+
 # 113 and 117 are our big ones
 # What happened then?
 country[c(113, 117), 'Date']
@@ -255,13 +258,22 @@ qqnorm(fit8$residuals)
 # This is looking much better.
 # Try it on the test set
 test9 = test[, finalvars]
+test99 = test9[-outindex1,]
+#outlier 114 118
+outindex1 = which(row.names(test9) %in% c('114', '113', '117','118'))
+
 fit9 = lm(I(1 / exchange_rate) ~ ., data=test9)
 summary(fit9)
-plot.lm(fit9, which=4)
+plot(fit9, which=4)
+
+#no outlier
+fit99 = lm(I(1 / exchange_rate) ~ ., data=test99)
+summary(fit99)
+plot(fit99, which=4)
+
 
 coef(fit8)
 coef(fit9)
-
 form10 = exchange_rate ~ gdp + interest_rate + inflation_KOR + inflation_USA
 fit10 = lm(form10, data= tdata8)
 # Do we still need to do inverse transform if outliers are gone? 
@@ -269,3 +281,106 @@ boxcox(fit10)
 # Still calls for the inverse transform even with the outliers removed.
 summary(fit10)
 
+add = lm(I(1/exchange_rate) ~ gdp + interest_rate + inflation_KOR + inflation_USA, data = alldata1)
+add1 = lm(govspending ~ gdp + interest_rate + inflation_KOR + inflation_USA, data = alldata1)
+
+pdf('Addedvariable.pdf')
+plot(add1$resi, add$resi, main = 'Added-variable Plot for govspending', xlab = 'e(govspending|otherX)', ylab = 'e(1/(exchangerate)|otherX)')
+dev.off()
+
+summary(add)
+summary(add1)
+
+alldata = rbind(test, tr)
+outlier = which(row.names(alldata) %in% c('114', '113', '117','118'))
+alldata1 = alldata[-outlier]
+
+fit11 = lm(I(1/exchange_rate) ~ gdp + interest_rate + inflation_KOR + inflation_USA, data = alldata)
+summary(fit11)
+anova(fit11)
+plot(fit11)
+coef(fit11)
+
+
+### scaling version ###
+sdata = sapply(alldata, scale)
+sdata = as.data.frame(sdata)
+
[…]

### 2014-12-14 paper.tex
sequence numbering, final draft

@@
 \[
-    Y = -0.69 c_1 \frac{G_{K}}{G_{US}}
-    -0.25 c_2 \frac{E_{K}}{E_{US}}
-    +0.59 c_3 \frac{R_{K}}{R_{US}}
-    +0.29 c_4I_{K} - 0.30 c_5 I_{US} + \epsilon.
+    Y = Intercept -c_1 \frac{G_{K}}{G_{US}}
+    -c_2 \frac{E_{K}}{E_{US}}
+    +c_3 \frac{R_{K}}{R_{US}}
+    +c_4I_{K} - c_5 I_{US} + \epsilon.
 \]
@@
 The 95\% confidence interval does not include 0 for any of these variables,
-so we are 95\% confident that these signs are correct. Since the predictor variables
+so we are 95\% confident that the signs of the coefficients are correct. Since the predictor variables
 are scaled or ratios they are unitless. Now the model has a
 natural interpretation. Suppose that the Korean economy grows while the US
-economy remains stagnant. Then the exchange rate should decrease, because
+economy remains stagnant. This should be associated with a decreasing
+exchange rate, because
 the Korean Won will have more value relative to the US Dollar.
@@
 
-Common sense and economic reasoning tells us that all of these should be
+Common sense and economic reasoning tells us that all of these statements should be
 true. The data and analysis confirm our intuition. Perhaps with the use of
@@
 
-Returning to out original questions, the results of this analysis suggest
-that if you want to exchange your currency
-to foreign currency, it would be better to do it when your home country's 
-GDP is good, or the exports are good, or the inflation is relatively lower.
+Now we return to our original questions. The results of this analysis 
+suggest a reasonable course of action if you want to exchange your currency
+to foreign currency. In relative terms, it is better to do it when your home country's 
+GDP is high, or the exports are high, or the inflation is lower.
 
@@
 
-\section{Appendix - Reproducibility}
-
-Above we mentioned that the data and all analysis in the report is self 
-updating. This is accomplished through the use of GNU \texttt{make}, which
-describes a DAG (Directed Acyclic Graph) of dependencies for the final
-report. \texttt{make} detects file modifications and will lazily run all
-commands and scripts required for the final output.
-
-What does this mean? Suppose that we need more variables to do the
-regression. Thanks to the use
-of industry standard ISO codes for the countries and proper parameters in
-our code base this becomes a simple task.
-We've stored string templates for the country level 
-variables in a file
-called \texttt{template.txt}. The data collection and preprocessing steps
-are automated, so we can add more variables simply by adding a row to
-\texttt{template.txt} and entering the command \texttt{make} from the shell
-prompt, which causes the following sequence of events:
-
-\begin{enumerate}
-    \item \texttt{make} detects the changed template file
-    \item The script \texttt{download.R} runs, updating the cache
-    \item The script \texttt{preprocess.R} runs, applying the appropriate
-        transformations and joining the tables to a form suitable for
-        analysis
-    \item All other scripts which depend on the preprocessed data run
-    \item The report output is produced
-\end{enumerate}
-
-In this case the output is a PDF, but it could just as easily be a web page
-on the department server, or an upload into another REST API.
-
-If we were doing this analysis for a client and need to run it again a year
-later with a new year of data 
-then all we have to do is type \texttt{make}. We'll still need to interpret
-plots and models, but the most time consuming work is done.
-
-One other aspect of reproducibility is version control. In this
-project we've used \texttt{Git} to collect snapshots of the project at
-every stage. The project is hosted publicly on
-Github\footnote{\url{https://github.com/clarkfitzg/stats206\_project}}.
-A look at the log file shows what work happened when and why,
-which is important for establishing provenance. In particular it shows 
-that we've been active on this project every day since
-December 1st, pushed more than 30 commits, and wrote nearly 3000 lines of
-code (including \LaTeX). Here's an example of a log entry:
-
-\begin{verbatim}
-commit e2beaf01b0a1457e7167989c1fae575c2a676f19
-Author: Clark Fitzgerald <clarkfitzg@gmail.com>
-Date:   Mon Dec 8 20:51:04 2014 -0800
-
-    implemented local caching, decoupled download step
-\end{verbatim}
-
-Through a little engineering the data analysis process
-can be made transparent, automated, extensible, and fully reproducible.
-
-\section{Appendix 1}
+\section{Appendix 1 - Plots}
 
@@
 \newpage
+.
+\newpage
 
-\section{Appendix 2}
+\section{Appendix 2 - R Output}
 
-\subsection{Model Selection Output}
+\subsection*{Model Selection Output}
 \begin{table}[ht]
@@
 
+\section{Appendix 3 - Reproducibility}
+
+Above we mentioned that the data and all analysis in the report is self 
+updating. This is accomplished through the use of GNU \texttt{make}, which
+describes a DAG (Directed Acyclic Graph) of dependencies for the final
+report. \texttt{make} detects file modifications and will lazily run all
+commands and scripts required for the final output.
+
+What does this mean? Suppose that we need more variables to do the
+regression. Thanks to the use
+of industry standard ISO codes for the countries and proper parameters in
+our code base this becomes a simple task.
+We've stored string templates for the country level 
+variables in a file
+called \texttt{template.txt}. The data collection and preprocessing steps
+are automated, so we can add more variables simply by adding a row to
+\texttt{template.txt} and entering the command \texttt{make} from the shell
+prompt, which causes the following sequence of events:
+
+\begin{enumerate}
+    \item \texttt{make} detects the changed template file
+    \item The script \texttt{download.R} runs, updating the cache
+    \item The script \texttt{preprocess.R} runs, applying the appropriate
+        transformations and joining the tables to a form suitable for
+        analysis
+    \item All other scripts which depend on the preprocessed data run
+    \item The report output is produced
+\end{enumerate}
+
+In this case the output is a PDF, but it could just as easily be a web page
+on the department server, or an upload into another REST API.
+
+If we were doing this analysis for a client and need to run it again a year
+later with a new year of data 
+then all we have to do is type \texttt{make}. We'll still need to interpret
+plots and models, but the most time consuming work is done.
+
+One other aspect of reproducibility is version control. In this
+project we've used \texttt{Git} to collect snapshots of the project at
+every stage. The project is hosted publicly on
+Github\footnote{\url{https://github.com/clarkfitzg/stats206\_project}}.
+A look at the log file shows what work happened when and why,
+which is important for establishing provenance. In particular it shows 
+that we've been active on this project every day since
+December 1st, pushed more than 30 commits, and wrote nearly 3000 lines of
+code (including \LaTeX). Here's an example of a log entry:
+
+\begin{verbatim}
+commit e2beaf01b0a1457e7167989c1fae575c2a676f19
+Author: Clark Fitzgerald <clarkfitzg@gmail.com>
+Date:   Mon Dec 8 20:51:04 2014 -0800
+
+    implemented local caching, decoupled download step
+\end{verbatim}
+
+Through a little engineering the data analysis process
+can be made transparent, automated, extensible, and fully reproducible.
+
 \end{document}
